### 2020-11-03 paper_old_analysis.tex

\documentclass[letterpaper,twocolumn,10pt]{article}
%\usepackage[letterpaper, total={6.5in, 9in}, columnsep=0.25in]{geometry}
\usepackage[letterpaper, total={7in, 9in}, columnsep=0.33in]{geometry}
\newcommand{\tronly}[2]{#1}
\usepackage{newtxtext}
\usepackage{appendix}
\tronly{}{
\usepackage{microtype}
\usepackage[leadingfraction=1,moderate]{savetrees}
}
\usepackage{comment}

\setlength{\parskip}{0pt}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
\setlength{\textfloatsep}{2pt plus 1.0pt minus 2.0pt}
\setlength{\floatsep}{2pt plus 1.0pt minus 2.0pt}
\usepackage{cite}
\usepackage{paralist}
\usepackage[hyphens]{url}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\setdefaultleftmargin{0em}{}{}{}{}{}
\setdefaultitem{{}}{}{}{}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}
%\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
%\do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
%\do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
%\do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
%\do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
%\do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
%\do\Y\do\Z}
\usepackage{caption}
% \captionsetup[figure]{font=small,labelfont=small,skip=3pt}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{thmtools, thm-restate}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{dsfont}
\usepackage{capt-of}
\usepackage{algorithmicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}
\usepackage[makeroom]{cancel}
\usepackage{datetime}
\renewcommand\ttdefault{cmtt}
\graphicspath{{./figures}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\let\oldemptyset\emptyset
\let\emptyset\varnothing
% New definitions
\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{assert}}
\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%⇧
\algnewcommand\algorithmiccontinue{\textbf{continue}}
\algnewcommand\algorithmicbreak{\textbf{break}}
\algnewcommand\Continue{\algorithmiccontinue}
\algnewcommand\Break{\algorithmicbreak}
\newcommand{\editremove}[1]{{\color{red}\sout{#1}}}
\newcommand{\editinsert}[1]{{\color{blue}#1}}
\newcommand{\editchange}[1]{{\color{orange}#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\setcounter{footnote}{1}
\newcommand{\Jon}[1]{{\color{blue} \textbf{Jon: } #1}}

\settimeformat{hhmmsstime}
\mmddyyyydate
\title{
\vspace{-0.30in}
    \Large \bf \editchange{Snowball}: The Power of Metastable Consensus}
%\tronly{{}
    \author{Team Rocket\footnote{\texttt{\ 0x\ 8a5d 2d32 e68b c500 36e4 d086 0446 17fe 4a0a 0296 b274 999b a568 ea92 da46 d533}
    }\\t-rocket@protonmail.com}
%}{
%\author{Paper \texttt{\#220}}
%}
%\tronly{\date{Revision: \today{} \currenttime{} UTC}}
\date{Revision: \today{} \currenttime{} UTC}
%{\date{\vspace{-2ex}}}
\usepackage[hang]{footmisc}
\setlength\footnotemargin{0em}
\begin{document}
\maketitle

\begin{abstract}
This paper introduces a new family of leaderless Byzantine fault tolerance protocols,
built around a metastable mechanism.
These protocols provide a strong probabilistic safety guarantee in the presence of Byzantine adversaries\tronly{, while their concurrent nature enables them to achieve high throughput and scalability.}{.}
Unlike blockchains that rely on proof-of-work, they are quiescent and green.
Unlike traditional consensus protocols having leader nodes that
process $O(n)$ information (in terms of number of bits exchanged), 
no node processes more than \editchange{$O(k\log n)$ or $O(k)$} for some security parameter~$k$.
% require $O(n^{2})$
% communication, their communication complexity ranges from
% $O(kn\log n)$
% to
% $O(kn)$
% for some security parameter $k \ll n$. \Jon{This sentence led me to assuem that the two ``best'' protocols were both going to be important; that there would be a reason to use the $O(kn\log n)$ version despite its poorer aymptotics. But when I got to later pages, I see you're only discussing Avalanche after all; the other algorithms are purely steps on the pedagogical garden path, right? Maybe remove the reference to the intermediate steps from the abstract?}
\tronly{% new paragraph

The paper describes the \editchange{Snowball protocol family}, analyzes its guarantees, and describes how it can be used to construct the core of an internet-scale electronic payment system.
}{}
\editinsert{A peer-to-peer payment system, Avalanche} is evaluated in a large scale deployment.
Experiments demonstrate that the system can achieve high throughput (3400 tps), provide low confirmation latency (1.35 sec), and scale well compared to existing systems that deliver similar functionality. For our implementation and setup, the bottleneck of the system is in transaction verification.
\end{abstract}

\section{Introduction}\tronly{}{\vspace{-0.5em}}

\tronly{Achieving agreement among a set of distributed hosts lies at the core of countless applications, ranging from Internet-scale services that serve billions of people~\cite{Burrows06,HuntKJR10} to cryptocurrencies worth billions of dollars~\cite{marketcapcryptocurrency}.}{}
To date, there have been two main families of solutions to this problem.
Traditional consensus protocols %, exemplified by PBFT~\cite{castro1999practical},
rely on all-to-all communication to ensure that all correct nodes reach the same decisions with absolute certainty.
Because they require quadratic communication overhead and accurate knowledge of membership, they have been difficult to scale
\tronly{to large numbers of participants.}{.}

On the other hand, Nakamoto consensus protocols~\cite{nakamoto2008bitcoin,GarayKL15, PassSS17, SompolinskyZ15, SompolinskyLZ16, SompolinskyZ18, BentovHMN17, EyalGSR16,Kokoris-KogiasJ16,PassS16a, PassS18} have become popular with the rise of Bitcoin.
These protocols provide a probabilistic safety guarantee: Nakamoto consensus decisions may revert with some probability $\varepsilon$.
A protocol parameter allows this probability to be rendered arbitrarily small\tronly{, enabling high-value financial systems to be constructed on this foundation.}{.}
This family is a natural fit for open, permissionless settings where any node can join the system at any time.
Yet, these protocols are costly, wasteful, and limited in performance.
By construction, they cannot quiesce: their security relies on constant participation by miners, even when there are no decisions to be made.
\tronly{Bitcoin currently consumes around 63.49 TWh/year~\cite{bitcoinpower}, about twice as all of Denmark~\cite{denmarkpower}.}{}
Moreover, these protocols suffer from an inherent scalability bottleneck that is difficult to overcome through simple reparameterization (see~\cite{CromanDEGJKMSSS16} for more details). % \Jon{I wanted the paper to expand on this point, but I missed it.}

This paper introduces a new family of consensus protocols.
Inspired by gossip algorithms, this family gains its safety through a deliberately metastable mechanism.
Specifically, the system operates by repeatedly sampling the network at random, and steering correct nodes towards the same outcome.
Analysis shows that this metastable mechanism is powerful: it can move a large network to an irreversible state quickly.
%{\color{red} We should clarify this.  The current statement is too strong.}
% though it is not always guaranteed to do so.
% \Jon{Not universal in that some applications require a hard guarantee?}

Similar to Nakamoto consensus, this new protocol family provides a probabilistic safety guarantee, using a tunable security parameter that can render the possibility of a consensus failure arbitrarily small.
Unlike Nakamoto consensus, the protocols are green, quiescent and efficient; they do not rely on proof-of-work~\cite{DworkN92} and do not consume energy when there are no decisions to be made.
The efficiency of the protocols stems from two sources: each node requires communication overheads ranging from $O(k \log n)$ to $O(k)$ for some small security parameter $k$ (whereas leader-based consensus protocols have one or more nodes that require $O(n)$ communication), and the protocols establish only a partial order among dependent transactions.

This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating values}. %conflicting transactions.

\editchange{Specifically, the theorecital, binary consensus protocol from the Snowball family guarantees liveness in logarithmic number of rounds in expectation. While, in the demonstrated payment system, Avalanche, the liveness is affected by the diversity of initial proposals and will be guaranteed when a unanimous proposal is given.
It guarantees liveness only for \emph{virtuous} transactions, i.e.\ those issued by correct clients and thus guaranteed not to conflict with other transactions.
However,} in a cryptocurrency setting, cryptographic signatures enforce that only a key owner is able to create a transaction that spends a particular coin.
Since correct clients follow the protocol as prescribed, they are guaranteed both safety and liveness.
In contrast, the protocols do not guarantee liveness for \emph{rogue} transactions, submitted by Byzantine clients, which conflict with one another.
Such decisions may stall in the network, but have no safety impact on virtuous transactions.
We show that this is a sensible tradeoff, and that resulting system is sufficient for building complex payment systems.

Overall, this paper makes one significant contribution: a brand new family of consensus protocols, based on randomized sampling and metastable decision. The protocols provide a strong probabilistic safety guarantee, and \editchange{the liveness guarantee suitable for building a decentralized, peer-to-peer, payment system.}

The next section provides the intuition behind the new protocols, followed by their full specification,
Section~\ref{sec:attack} briefly discusses typical attack vectors and their possible defense,
Section~\ref{sec:analysis} provides methodology used
by our formal analysis of safety and liveness in Appendix~\ref{sec:full-analysis},
%Section~\ref{sec:implementation} describes a Bitcoin-like payment system,  Section~\ref{sec:evaluation} evaluates the system,
Section~\ref{sec:implementation} describes \editchange{Avalanche, a Bitcoin-like payment system,
Section~\ref{sec:evaluation} evaluates Avalanche},
Section~\ref{sec:related-work} presents related work, and finally, Section~\ref{sec:conclusions} summarizes our contributions.

\section{Approach}\tronly{}{\vspace{-0.5em}}
\label{sec:approach}
We start with a non-Byzantine protocol, Slush, and progressively build up \editchange{Snowflake and Snowball}, all based on the same common metastable mechanism.
\tronly{
\editremove{Slush, Snowflake, and Snowball} \editinsert{They} are single-decree binary consensus protocols
of increasing robustness\editremove{---Avalanche extends Snowball into a full
cryptocurrency solution; other extensions for other applications are certainly
possible}.
We provide full specifications for the protocols in this section, and defer the analysis to the next section, and present formal
proofs in the appendix.
}{}

\editremove{Overall, this protocol family achieves its properties by humbly cheating in three different ways.}
\editremove{First, taking} \editinsert{Taking} inspiration from Bitcoin, we adopt a safety guarantee that is probabilistic.
\tronly{This probabilistic guarantee is indistinguishable from traditional safety guarantees in practice, since appropriately small choices of $\varepsilon$ can render consensus failure practically infeasible, less frequent than CPU miscomputations or hash collisions.}{}

\editinsert{On the other hand, our practical payment system, Avalanche (Section~\ref{sec:implementation}), achieves its
properties by humbly cheating in two ways.}
\editremove{Second, }\editinsert{First, }instead of a single replicated state machine (RSM) model, where the system determines a sequence of totally-ordered transactions $T_0, T_1, T_2, \ldots$ issued by any client, we adopt a \emph{parallel consensus model}, where each client interacts independently with its own RSMs and optionally transfers ownership of its RSM to another client. (MOVED to Model Section) The system establishes only a partial order between dependent transactions.
\editremove{Finally, }\editinsert{Second, }for Avalanche, we provide no liveness guarantee for misbehaving clients, but ensure that well-behaved clients will eventually be serviced.
These techniques, in conjunction, enable the system to implement a useful Bitcoin-like cryptocurrency, with drastically better performance and scalability.

\editchange{Because our Snowball family provides strong safety guarantee and liveness guarantee},
it is possible to build other applications involving large-scale probabilistic consensus. We focus on a cryptocurrency application because of many challenges it poses.
%Nevertheless, Snowball does provide liveness guarantee for conflicting values, so it is possible to extend this single-decree protocol for other possible applications involving large-scale probabilistic consensus.
%\Jon{Are there other applications where this consensus model would be valuable? It's not a BFT RSM, but it would be very intriguing if there were other additional applications.}

\subsection{Model, Goals, Threat Model}\tronly{}{\vspace{-0.5em}}

We assume a collection of nodes, $\mathcal{N}$,
composed of correct nodes $\mathcal{C}$ and
Byzantine nodes $\mathcal{B}$, where $n = |\mathcal{N}|$.

\editinsert{In Avalanche}, we adopt what is commonly known as
Bitcoin's unspent transaction output (UTXO) model.
In this model,
\emph{clients} are authenticated and issue cryptographically signed
transactions that fully consume an existing UTXO and issue new UTXOs.
Unlike nodes, clients do not participate in the decision process, but only
supply transactions to the \emph{nodes} running the consensus protocol.
Two transactions \emph{conflict} if
they consume the same UTXO and yield different outputs.
Correct clients never issue conflicting transactions, nor is it possible
for Byzantine clients to forge conflicts with transactions
issued by correct clients.
However, Byzantine clients can issue multiple transactions that conflict
with one another, and correct clients can only consume one of
those transactions.
\editchange{To make use of Snowball consensus in Avalanche}, then, is to \emph{accept} a
set of non-conflicting transactions in the presence of Byzantine behavior.
Each client can be considered a replicated state machine whose
transitions are defined by a totally ordered list of accepted transactions.

Our family of protocols provide the following guarantees with high probability (whp):
\begin{compactitem}
% RVR: this property is unnecessary because the liveness property already ensures non-triviality
%    \item \textbf{Non-triviality}: To ensure that the protocol carries out real work, decisions must include values proposed by external clients;

    \item \textbf{P1. Safety.} No two correct nodes will accept conflicting transactions.
% the following is an assumption, not a property.
%	for some well-characterized $|\mathcal{B}|$.

    \item \textbf{P2. Liveness.} Any transaction issued by a correct client (aka virtuous transaction) will eventually be accepted by every correct node.
\end{compactitem}

Instead of unconditional agreement, our approach guarantees that safety will be upheld with probability $1-\varepsilon$, where the choice of the security parameter $\varepsilon$ is under the control of the system designer and applications.

We assume a powerful adaptive adversary capable of observing the internal state and communications of every node in the network, but not capable of interfering with communication between correct nodes.
Our analysis assumes a synchronous network, while our deployment and evaluation is performed in a partially synchronous setting.
We conjecture that our results hold in partially synchronous networks, but the proof is left to future work.
We do not assume that all members of the network are known to all participants, but rather may temporarily have some discrepancies in network view.
We assume a safe bootstrapping system, similar to that of Bitcoin, that enables a node to connect with sufficiently many correct nodes to acquire a statistically unbiased view of the network.
We do not assume a PKI\@.
We make standard cryptographic assumptions related to public key signatures and hash functions.

\subsection{Slush: Introducing Metastability}\tronly{}{\vspace{-0.5em}}
The core of our approach is a single-decree consensus protocol,
inspired by epidemic or gossip protocols.
% In this approach, a state in which all possible outcomes are equally
% probable is metastable and disturbances are likely to cause the
% protocol to converge to one of multiple stable states.
% We will call such consensus protocols metastable themselves.
%We can model a preference for a decision as a kind of infection: a node catches a disease, and converts other nodes to carry the same disease in epidemic fashion.
%If a threshold of its interactions are with other infected hosts, it flips its color to the competing agent and joins their ranks.
%
The simplest metastable protocol, Slush, is the foundation of this family, shown in Figure~\ref{fig:slush-loop}.
Slush is \emph{not} tolerant to Byzantine faults, but serves as an illustration for the BFT protocols that follow.
For ease of exposition, we will describe the operation of Slush using a decision between two conflicting colors, red and blue.
% , though the technique generalizes to any number of competing proposed values.
% Slush operates as follows.

In Slush, a node starts out initially in an uncolored state.
Upon receiving a transaction from a client, an uncolored node updates its own color to the one carried in the transaction and initiates a query.
To perform a query, a node picks a small, constant sized ($k$) sample of the network uniformly at random, and sends a query message.
Upon receiving a query, an uncolored node adopts the color in the query, responds with that color, and initiates its own query, whereas a colored node simply responds with its current color.
If $k$ responses are not received within a time bound, the node picks an additional sample from the remaining nodes uniformly at random and queries them until it collects all responses.
Once the querying node collects $k$ responses, it checks if a fraction $\ge\alpha k$ are for the same color, where $\alpha > 0.5$ is a protocol parameter.
If the $\alpha k$ threshold is met and the sampled color differs from the node's own color, the node flips to that color.
It then goes back to the query step, and initiates a subsequent round of query, for a total of $m$ rounds.
Finally, the node decides the color it ended up with at time $m$.

\algnewcommand{\IIf}[1]{\State\algorithmicif\ #1\ \algorithmicthen}
\algnewcommand{\EndIIf}{\unskip\ \algorithmicend\ \algorithmicif}

\newcommand{\codecolor}{\textrm{col}}
\newcommand{\codecount}{\textrm{cnt}}
\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{onQuery}{$v, \codecolor'$}
        \IIf{$\codecolor = \bot$} $\codecolor \coloneqq \codecolor'$
        \State\Call{respond}{$v, \codecolor$}
    \EndProcedure
    \Procedure{slushLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \coloneqq \codecolor_0$ \textrm{// initialize with a color}
        \For{$r \in \{1\ldots m\}$}
            \State \textrm{// if $\bot$, skip until \textsc{onQuery} sets the color}
            \If{$\codecolor = \bot$}
                \Continue
            \EndIf
            \State \textrm{// randomly sample from the known nodes}
            \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
            \State $P \coloneqq \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
                \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                    \State $\codecolor \coloneqq \codecolor'$
                \EndIf
            \EndFor
        \EndFor
        \State \Call{accept}{$\codecolor$}
    \EndProcedure
    \captionof{figure}{Slush protocol. Timeouts elided for readability.}\label{fig:slush-loop}
\end{algorithmic}
\end{figure}

This simple protocol has some curious properties.
First, it is almost \emph{memoryless}: a node retains no state between rounds other than its current color, and in particular maintains no history of interactions with other peers.
Second, unlike traditional consensus protocols that query every participant, every round involves sampling just a small, constant-sized slice of the network at random.
\tronly{Third, even if the network starts in the metastable state of a 50/50 red-blue split, random perturbations in sampling will cause one color to gain a slight edge and repeated samplings afterwards will build upon and amplify that imbalance.}{}
Finally, if $m$ is chosen high enough, Slush ensures that all nodes will be colored identically whp.
Each node has a constant, predictable communication overhead per round, and $m$ grows logarithmically with $n$. %XXX where do we do this?

\tronly{
%If Slush is deployed in a network with Byzantine nodes, the adversary can interfere with decisions.
%In particular, if the correct nodes develop a preference for one color, the adversary can attempt to flip nodes to the opposite so as to keep the network in balance.
%The Slush protocol lends itself to analysis but does not, by itself, provide a strong safety guarantee in the presence of Byzantine nodes, because the nodes lack state.
%We address this in our first BFT protocol.
% Ittay:
The Slush protocol does not provide a strong safety guarantee in the presence of Byzantine nodes.
In particular, if the correct nodes develop a preference for one color, a Byzantine adversary can attempt to flip nodes
to the opposite so as to keep the network in balance, preventing a decision.
We address this in our first BFT protocol that introduces more state storage at the nodes.
}{
We next examine how to extend Slush to tolerate Byzantine behavior.
}

\subsection{Snowflake: BFT}\tronly{}{\vspace{-0.5em}}

Snowflake augments Slush with a single counter that captures the strength of a node's conviction in its current color.
This per-node counter stores how many consecutive samples of the network by that node have all yielded the same color.
A node accepts the current color when its counter exceeds $\beta$, another security parameter.
Figure~\ref{fig:snowflake-loop} shows the amended protocol, which includes
the following modifications:

\begin{compactenum}
	\item Each node maintains a counter $\mathit{cnt}$;
    \item Upon every color change, the node resets $\mathit{cnt}$ to 0;
    \item Upon every successful query that yields $\ge \alpha k$ responses for the same color as the node, the node increments $\mathit{cnt}$.
\end{compactenum}

\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{snowflakeLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \coloneqq \codecolor_0$, $\codecount \coloneqq 0$
        \While{\textrm{undecided}}
            \If{$\codecolor = \bot$}
                \Continue
            \EndIf
            \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
            \State $P \coloneqq \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                \If{$\codecolor' \neq \codecolor$}
                    \State $\codecolor \coloneqq \codecolor'$, $\codecount \coloneqq 0$
                \Else
                    \IIf {$\texttt{++}\codecount > \beta$} \Call{accept}{$\codecolor$}
                \EndIf
            \EndIf
            \EndFor
        \EndWhile
    \EndProcedure
    \captionof{figure}{Snowflake.}\label{fig:snowflake-loop}
\end{algorithmic}
\end{figure}

When the protocol is correctly parameterized for a given threshold of Byzantine nodes and a desired $\varepsilon$-guarantee, it can ensure both safety (P1) and liveness (P2).
As we later show, there exists a phase-shift point after which correct nodes are more likely to tend towards a decision than a bivalent state.
Further, there exists a point-of-no-return after which a decision is inevitable.
The Byzantine nodes lose control past the phase shift, and the correct nodes begin to commit past the point-of-no-return, to adopt the same color, whp.

\tronly{
\begin{figure}
    \small
\begin{algorithmic}[1]
    \Procedure{snowballLoop}{$u, \codecolor_0 \in \{\texttt{R}, \texttt{B}, \bot\}$}
        \State $\codecolor \coloneqq \codecolor_0$, $\textrm{lastcol} \coloneqq \codecolor_0$, $\codecount \coloneqq 0$
        \State $d[\texttt{R}] \coloneqq 0$, $d[\texttt{B}] \coloneqq 0$
        \While{\textrm{undecided}}
            \IIf{$\codecolor = \bot$} \Continue
            \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
            \State $P \coloneqq \texttt{[}\Call{query}{v, \codecolor}\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
            \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                \State $d[\codecolor']\texttt{++}$
                \If {$d[\codecolor'] > d[\codecolor]$}
                        \State$\codecolor \coloneqq \codecolor'$
                \EndIf
                \If{$\codecolor' \neq \textrm{lastcol}$}
                    \State$\textrm{lastcol} \coloneqq \codecolor'$, $\codecount \coloneqq 0$
                \Else
                    \IIf {$\texttt{++}\codecount > \beta$} \Call{accept}{$\codecolor$}
                \EndIf
            \EndIf
            \EndFor
        \EndWhile
    \EndProcedure
    \captionof{figure}{Snowball.}\label{fig:snowball-loop}
\end{algorithmic}
\end{figure}
}{}

\subsection{Snowball: Adding Confidence}\tronly{}{\vspace{-0.5em}}

Snowflake's notion of state is ephemeral: the counter gets reset with every color flip.
While, theoretically, the protocol is able to make strong guarantees with
minimal state, we will now improve the protocol to make it harder to
attack by incorporating a more permanent notion of belief.
Snowball augments Snowflake with \emph{confidence counters} that capture the number of queries that have yielded a threshold result for their corresponding color (Figure ~\ref{fig:snowball-loop}).
A node decides if it gets $\beta$ consecutive chits for a color. However, it only changes preference based on the total accrued confidence.
The differences between Snowflake and Snowball are as follows:

\begin{compactenum}
    \item Upon every successful query, the node increments its confidence counter for that color.
    \item A node switches colors when the confidence in its current color becomes lower than the confidence value of the new color.
\end{compactenum}

Snowball is not only harder to attack than Snowflake, but is more easily
generalized to multi-decree protocols.

\editremove{MOVED Avalanche to the implementation section}
\section{Analysis}
\label{sec:analysis}
Due to space limits, we move the full safety analysis to Appendix~\ref{sec:full-analysis}.
There, we show that under certain independent and distinct assumptions, the \editchange{Snowball} family of consensus protocols provide properties P1 (safety) and P2 (liveness) with probability $1 - \varepsilon$, where $\varepsilon$ is a security parameter that can be set according to the needs of the system designer.
In this section, we summarize the methodology used for our analysis and provide some insight into our core results.
\subsection{Safety}
% Analyzing the protocols introduced in this paper
%differ from classical as well as Nakamoto consensus-style protocols. 
%Even simple, non-linear dynamical systems often incur complicated analyses as they may not admit analytical solutions.
% In particular, since the state of a node depends on randomly sampling other nodes, possibly inducing nondeterministic dependency cycles, analyzing all possible state executions is computationally infeasible. 
% Here, we employ stochastic techniques used in the analysis of epidemic networks, which may also be of independent interest to the distributed systems community. 
%Here, we employ 
% requires complex stochastic techniques to reason about the state of the network in aggregate.

\paragraph{Scheduler, Network, and Adversary.} 
We analyze the protocols through a global view. 
Like Bitcoin and other blockchain protocols based on gossip dissemination, we assume that gossip spreads to all correct nodes within a known time bound.  
In other words, we assume a synchronous communication network, where the system makes progress in rounds, and where at the beginning of each round the system scheduler chooses a single correct node $u$ uniformly at random. 
Node $u$ will sample and update its state by the end of the round. 
The Byzantine adversary will be aware of the identity of $u$. Furthermore, the adversary has full knowledge of the internal state of all nodes in the network at all times, and is able to fully update the state of all nodes under its control immediately after $u$ chooses its neighbor sample set. 
In essence, the adversary is only constrained by an inability to directly update the state of correct nodes.

\paragraph{System Dynamics.} We can think of the network as a set of nodes either colored red or blue. Each new round will update the state of one of the nodes, either changing its color, or increasing its confidence in its current color. 
The dynamics of the system resemble those of epidemic networks, where nodes update their state based on the state of their neighbors, using a function chosen over some probability distribution. 
It would be infeasible to keep track of all possible execution paths, since at every round, the possible number of branching executions is equal to the number of possible correct nodes times all possible $k$-sample choices of outcomes that the node may sample. Furthermore, the difficulty of this task is greatly amplified due to the Byzantine adversary, whose optimal attack strategy is chosen over an unknown function, possibly itself nondeterministic.

The simplification that allows us to analyze this system is to obviate the need to keep track of all of the execution paths, as well as all possible adversarial strategies, and rather focus entirely on a single state of interest, without regards to how we achieve this state. 
More specifically, the core extractable insight of our analysis is in identifying the \textit{irreversibility} state of the system, the state upon which so many correct nodes have usurped either red or blue that reverting back to the minority color is highly unlikely.

\paragraph{Slush}
We model the Slush protocol is through a birth-death Markov process with two absorbing states, where either all nodes are red or all nodes are blue. 
In our construction, all states (excluding absorbing states) are transient, meaning that there is a nonzero probability of never returning to that state. 
Our analysis shows that the Slush protocol reaches, in finite time, an absorbing state with non-zero probability.
Furthermore, the probability of convergence to one of the two colors can be precisely measured using only a local round counter. In fact, Slush converges to a decision in close to logarithmic steps whp.

% Another interesting outcome of the protocol is that optimal convergence rate occurs at $k = \alpha = 1$, meaning that the network will reach unanimous decision fastest when every node simply polls a single other neighbor. 
% As we discuss for later protocols, however, there is a tension between convergence rate and safety guarantees. Smaller values of $k$ impede the ability of the correct nodes to protect against an adversarial attack.

\paragraph{Snowflake}
For Snowflake, we relax the assumption that all nodes are correct and assume that some fraction of nodes are adversarial.
An important insight is that there exists an irreversible state, or \textit{point of no return}, after which the system will converge to an absorbing state whp. Furthermore a correct node only decides when the system is beyond the point of no return. Composing these two guarantees together, the probability of a safety violation is strictly less than $\epsilon$, which can be configured as desired. 
Unsurprisingly, there is an inherent tension between safety and liveness, but suitable parameters can be found that are practical.
Larger values of $k$ obtain higher levels of security for correct nodes, at the expense of slower convergence.

\paragraph{Snowball}
Snowball is an improvement over Snowflake, where random perturbations in network samples are reduced by introducing a partial form of history, which we refer to as confidence. 
Modeling Snowball with a Markov chain is difficult because of a state space explosion problem. In particular, it is not sufficient to simply keep track of color preferences of each correct node, the analysis must also maintain information about their confidence. 
To make analysis possible, we structure the mathematical foundations via a game of balls and urns, where each urn represents one of the correct nodes, and the ball count correspond to confidences in either color. 
Using this model, the analysis applies various tail inequalities to prove that the safety guarantees of Snowball are strictly stronger than those of Snowflake. In particular, once the system has reached the point of no return, the probability of reverting is strictly lower than in Snowflake.

\paragraph{Avalanche}
\editremove{%
MOVED to implementation}

\subsection{Liveness}\tronly{}{\vspace{-0.5em}}
% \tronly{
% \noindent\textbf{Slush.}
% Slush is a non-BFT protocol, which terminates whp within a finite number of rounds.
% }{}

\noindent\textbf{Snowflake \& Snowball.}
Both protocols make use of a counter to keep track of consecutive majority support. Since the adversary is unable to forge
a conflict for a virtuous transaction, initially,
all correct nodes will have color red or
$\bot$. A Byzantine node cannot respond to any query with any answer
other than red since it is unable to forge conflicts and $\bot$ is not allowed by protocol. Therefore, the only misbehavior for the Byzantine node is refusing to answer. Since the correct node will re-sample if the query times out, by expected convergence in \tronly{Equation~\ref{equation:expected_convergence}}{\cite{avalanche}} of Appendix~\ref{sec:full-analysis},
 all correct nodes will terminate with the unanimous red value within a finite number of rounds whp.

\editremove{%
\noindent\textbf{Avalanche.}
MOVED to implementation}

\subsection{Additional Insights}
Two interesting insights arise from our analysis.
First, the protocols discussed in this work lead to both safety and liveness guarantees whose underlying function is smooth, rather than a step function. 
In many other consensus protocols, safety is guaranteed with up to a fixed threshold number (e.g.~1/3) of adversarial nodes, beyond which no guarantees are provided. In our protocols, the guarantees degrade gracefully with an adversarial percentage beyond the pre-established bound. For example, optimal system parameters can be chosen to tolerate precisely 1/5 adversarial presence with failure probability $\epsilon$. However, if the system faces an adversarial presence greater than 1/5, then the probability of failure degrades to slightly above $\epsilon$, rather than immediately to $1$. 
Second, these protocols externalize the safety and liveness tradeoffs. 
The system designer may choose to guarantee safety even under catastrophic events, such as an adversarial presence beyond $1/2$, at the expense of liveness.  
This presents a powerful knob not available in classical or Nakamoto-based consensus protocols.

% \begin{compactitem}
% \item\textbf{Eventually good ancestry.} Virtuous transactions can be retried by picking new parents, selected from a set that is more likely to be preferred.
% Ultimately, one can always attach a transaction to decided parents to completely mitigate this risk.
% \Jon{So why don't we always do this? Later the paper clarifies that moving up the chain reduces the efficiency
%     and increases latency for the proposed transaction. I got pretty lost in the last two paragraphs of section 4}
% A simple technique for parent selection is to select new parents for a virtuous transaction at successively lower heights in the DAG, proceeding towards the genesis vertex.
% This procedure is guaranteed to terminate with uncontested, decided parents, ensuring that the transaction cannot suffer liveness failure due to rogue transactions.

% \item\textbf{Sufficient chits.} A secondary mechanism is necessary to ensure that virtuous transactions with decided ancestry will receive sufficient chits.
% To ensure this, correct nodes examine the DAG for virtuous non-nop transactions that lack sufficient progeny and emit nop transactions to help increase their confidence.
% A nop transaction has just one parent and no application side-effects, and can be issued by any node. They cannot be abused by Byzantine nodes because, even though nops trigger new queries, they do not automatically grant chits.

% With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.
% \end{compactitem}

%XXX discuss how no-ops can be implemented

\subsection{Attack Vectors}
\label{sec:attack}
There are a multitude of attack vectors against consensus protocols.  
We consider the two most important ones.
\paragraph{Sybil Attacks}
Consensus protocols provide their guarantees based on assumptions that only a fraction of participants are adversarial.
These bounds could be violated if the network is naively left open to arbitrary participants.
In particular, a Sybil attack~\cite{douceur2002sybil}, wherein a large number of identities are generated by an adversary, could be used to exceed the adversarial bound.

A long line of work, including PBFT~\cite{castro1999practical}, treats the Sybil problem separately from consensus, and rightfully so, as Sybil control mechanisms are distinct from the underlying, more complex agreement protocol\footnote{This is not to imply that every consensus protocol can be coupled/decoupled with every Sybil control mechanism.}.
Nakamoto consensus, for instance, uses proof-of-work~\cite{aspnes2005exposing} to limit Sybils, which requires miners to continuously stake a hardware investment.
Other protocols, discussed in Section~\ref{sec:related-work}, rely on proof-of-stake or proof-of-authority.
The consensus protocols presented in this paper can adopt any
% established
Sybil control mechanism, although proof-of-stake is most aligned with their quiescent operation.
The full design of a cryptocurrency incorporating staking, unstaking and minting mechanism is beyond the scope of this paper, whose focus is on the core consensus protocol.

% Sybil control mechanisms are orthogonal and separate from the consensus protocols. All such protocols, excluding Nakamoto-style consensus protocols, require a mechanism for network identity establishment. In a real implementation of Avalanche, the Sybil problem would be solved through a staking mechanism.

\paragraph{Flooding Attacks}
Flooding/spam attacks are a problem for any distributed system. 
Without a protection mechanism, an attacker can generate large numbers of transactions and flood the DAG, consuming storage. 
% Although there is no proper, protocol-level methodology to prevent such an attack, there are very effective heuristics that we can employ. 
There are a multitude of techniques to deter such attacks, including network-layer protections, proof-of-authority, or economic mechanisms. 
In Avalanche, we use transaction fees, making such attacks costly even if the attacker is sending money back to addresses under its control.

\subsection{Communication Complexity}\tronly{}{\vspace{-0.5em}}
Since liveness is not guaranteed for rogue transactions, we focus our message complexity analysis solely for the case of virtuous transactions.
For the case of virtuous transactions, Snowflake and Snowball are both guaranteed to terminate after $O(kn\log n)$ messages.
This follows from the well-known results related to epidemic algorithms\cite{demers1987epidemic}, and is confirmed by Table~\ref{table:growth_worst_case} in Appendix~\ref{sec:full-analysis}.

\editremove{Avalanche: MOVED to implementation}

\section{Peer-to-Peer Payment System}\tronly{}{\vspace{-0.5em}}
\label{sec:implementation}

%\label{sec:evaluation}

\editchange{We have implemented a bare-bones payment system, Avalanche, and
fully ported Bitcoin transactions to it.} In this section, \editchange{we describe the design and sketch how the implementation can support the value transfer primitive at the center of cryptocurrencies.} \editremove{and then examine its throughput, scalability, and latency through a large scale deployment on Amazon AWS\@, and finally, provide a comparison to known results from other systems.}

Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
and inflation control. While we have solutions for these issues, their full discussion is beyond the scope of this paper.
%In this section,
%we focus on how Avalanche can support the value transfer primitive at the center of cryptocurrencies.

\input{avalanche}

\editremove{\noindent\textbf{UTXO Transactions.}}
\editinsert{\subsection{Support UTXO Transactions}}
In addition to the DAG structure in Avalanche, a UTXO graph that captures
spending dependency is used to realize the ledger for the payment system. To
avoid ambiguity, we denote the transactions that encode the data for money
transfer \emph{transactions}, while we call the
transactions ($T \in \mathcal{T}$) in Avalanche's DAG \emph{vertices}.

%\tronly
%{
%Each transaction represents a money transfer that takes several inputs from
%source accounts and generates several outputs to destinations. As a UTXO-based system
%that keeps a decentralized ledger, balances are kept by the \emph{unspent outputs}
%of transactions.
%
%More specifically, a transaction $\texttt{TX}_a$ maintains
%a list of inputs: $\texttt{In}_{a1}$, $\texttt{In}_{a2}$, $\cdots$. Each input
%has two fields: the reference to an unspent transaction output and a spend script.
%The unspent transaction output uniquely refers to an output of a previously made transaction.
%The script snippet will be prepended to the script from the referred output forming a complete computation.
%It could typically be a cryptographic proof of validity, but could also be any Turing-complete
%computation in general.
%Each output $\texttt{Out}_{a1}$, $\texttt{Out}_{a2}$, $\cdots$,
%contains some amount of money and a script which typically contains cryptographic verification that takes the proof from the future input and verifies validity.
%
%In our payment system, there are \emph{addresses} representing different
%accounts by cryptographic keys. The public
%key is used as the identity for recipients in the output scripts, while
%the private key is for creating signatures in the
%input scripts, spending the available funds. Only the key owner is able spend the unspent output by creating an input with
%the signature in a new transaction.
%
%Due to the possibility of double spending by the private key owner,
%cryptocurrencies such as Bitcoin use a blockchain as the linear log to reject the
%transaction that comes later in the log and tries to spend some output twice.
%Instead, in our payment system, we use Avalanche to resolve the double-spend
%conflicts in each conflict set, without maintaining a linear log.
%
%If we could assume each transaction can only have one single input, the
%initialization of Avalanche would be straightforward. We let each vertex on the
%underlying DAG be one transaction.  The conflict set in Avalanche is the set of
%transactions that try to spend the same unspent output. The conflict sets are
%disjoint because each transaction only has one input spending one
%unspent output, and thus belongs to exactly one set.
%
%%If we support transactions with multiple inputs using the same
%%method as in single-input case, we will violate the disjoint conflict set model assumed by
%%Avalanche, because the double spending relation between transactions with multiple
%%inputs is not transitive: $\texttt{TX}_a$ may double spend the same output with
%%$\texttt{TX}_b$ due to the same input value $\texttt{In}_{a1} =
%%\texttt{In}_{b1}$, and $\texttt{TX}_b$ might double spend with $\texttt{TX}_c$
%%due to $\texttt{In}_{b2} = \texttt{In}_{c1}$, but $\texttt{TX}_a$ may not
%%double spend with $\texttt{TX}_c$ at all.
%
%Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
%To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as an Avalanche vertex,
%and use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted.
%Since the acceptance for each pair is meaningful for
%the payment system only if all pairs from the same transaction are accepted, we can tie the fate of these
%pairs from the same transaction together by a single, bundled query: the queried
%node will only answer ``yes'' if all of the pairs are strongly preferred
%according to the DAG\@. This more conservative predicate will not undermine safety because merely introducing transactions that gather no chits will not increase confidence value in the
%protocol.
%
%\begin{figure}[t]
%\begin{center}
%    \input{figures/cash-system-a}
%    \captionof{figure}{The actual DAG implementation at transition granularity.}
%    \label{fig:cash-system-a}
%\end{center}
%\end{figure}
%
%Figure~\ref{fig:cash-system-a} demonstrates the actual implementation where the DAG is built at transaction granularity, whereas Figure~\ref{fig:cash-system-b} shows the equivalent logic of the underlying protocol, where vertices are at transaction-input granularity.
%}
%{
We inherit the transaction and address mechanisms from Bitcoin. At their simplest, transactions consist of multiple inputs and outputs, with corresponding redeem scripts.
Addresses are identified by the hash of their public keys, and signatures are generated by corresponding private keys.
The full scripting language is used to ensure that a redeem script is authenticated to spend a UTXO\@.
UTXOs are fully consumed by a valid transaction, and may generate new UTXOs spendable by named recipients.
Multi-input transactions consume multiple UTXOs, and in Avalanche, may appear in multiple conflict sets.
To account for these correctly, we represent \emph{transaction-input} pairs (e.g. $\texttt{In}_{a1}$) as Avalanche vertices.
The conflict relation of transaction-input pairs are transitive because of each pair only spends one unspent output.
Then, we use the conjunction of \textsc{isAccepted} for all inputs of a transaction to ensure that no transaction will be accepted unless all its inputs are accepted (Figure~\ref{fig:cash-system-b}). 
Following this idea, we finally implement the DAG of transaction-input pairs such that multiple
transactions can be batched together per query.

%}

%However, supporting only a single input is both inconvenient and inefficient.
%One has to issue multiple transactions to merge the unspent balance in order to
%make larger payments. Moreover, with more than one output, the growth of unspent output
%set (i.e. UTXO set) size could be exponential, causing
%trouble for maintenance. However, if we also require a single output for
%all transactions, the granularity of coins cannot be changed, and one
%separated transaction is needed for transfering each unit of the currency.

\begin{figure}[t]
\begin{center}
    \input{figures/cash-system-b}
    \captionof{figure}{The underlying logical DAG structure used by Avalanche.
    The tiny squares with shades are dummy vertices which just help form the
    DAG topology for the purpose of clarity, and can be replaced by direct
    edges. The rounded gray regions are the conflict sets.}\label{fig:cash-system-b}
\end{center}
\end{figure}

%\noindent\textbf{Parent Selection.}
%The goal of the parent selection algorithm is to yield a well-structured DAG
%that maximizes the likelihood that virtuous transactions will be quickly
%accepted by the network. While this algorithm does not affect the safety of the
%protocol, it affects liveness and plays a crucial role in determining the shape of the
%DAG\@. A good parent selection algorithm grows the DAG
%in depth with a roughly steady ``width.'' The DAG should not diverge like a
%tree or converge to a chain, but instead should provide concurrency so nodes can work on multiple fronts.
%
%There are inherent trade-offs in the parent selection algorithm: selecting well-accepted parents makes it
%more likely for a transaction to find support, but can lead to vote dilution. Further, selecting more
%recent parents at the frontier of the DAG can lead to stuck transactions, as the parents may turn out
%to be rogue and remain unsupported.  In the following discussion, we illustrate this dilemma.
%We assume that every transaction will select a small number, $p$, of parents.
%We focus on the selection of eligible parent set, from which a subset of size $p$ can be chosen arbitrarily.
%%\Jon{Why is it important to choose parents randomly? What's the attack?}
%
%\tronly{
%Perhaps the simplest idea is to issue a fresh transaction with parents picked uniformly at random
%among those transactions that are currently strongly preferred. Specifically, we can adopt the predicate
%used in the voting rule to determine eligible parents on which a node would vote positively, as follows:
%\[
%    \mathcal{E} = \{T: \forall T\in\mathcal{T}, \textsc{isStronglyPreferred}(T)\}.
%\]
%
%But this strategy will yield large sets of eligible parents, consisting mostly of historical, old transactions.
%When a node samples the transactions uniformly from $\mathcal{E}$, the resulting DAG
%will have large, ever-increasing fan-out. Because new transactions will have scarce progenies,
%the voting process will take a long time to build the required confidence on any given new transaction.
%
%In contrast, efforts to reduce fan-out and control the shape of the DAG by selecting the recent transactions at
%the decision frontier suffer from another problem. Most recent transactions will have very low confidence,
%simply because they do not have enough descendants. Further, their conflict sets may not be well-distributed
%and well-known across the network, leading to a parent attachment under a transaction that will never be supported.
%This means the best parent candidates lie somewhere near the frontier, but not too far deep in history.
%
%The adaptive parent selection algorithm chooses parents by starting at the DAG frontier and retreating towards the
%genesis vertex until finding an eligible parent.
%}{
%The frontier transactions are those that currently do not have any children in the node's
%$\mathcal{T}$.  The set of strongly preferred transactions $\mathcal{E}$ is
%filtered by an additional criterion that the confidence should be larger than
%zero if the transaction has known conflicts. This filters out 0-confidence transactions from
%unstable conflict sets.  If there exist $p$ transactions
%in the filtered $\mathcal{E}$, then the adaptive selection will yield these transactions.
%
%Otherwise, the algorithm tries the parents of the transactions in $\mathcal{E}$,
%thus increasing the chance of finding more stabilized transactions as it
%retreats.
%}
%%XXX issue no-ops after a timeout
%The retreating search is guaranteed to terminate when it reaches the genesis
%vertex.  Formally, the selected parents in this adaptive selection algorithm
%is:
%
%\begin{center}
%\begin{algorithmic}[1]
%    \small
%    \Function{parentSelection}{$\mathcal{T}$}
%    \tronly{}{\State $\mathcal{E} = \{T: \forall T\in\mathcal{T}, \textsc{isStronglyPreferred}(T)\}$.}
%        \State $\mathcal{E}' \coloneqq \{T: \forall T\in\mathcal{E}, |\mathcal{P}_T| = 1 \lor d(T) > 0\}$.
%        \State \Return $\{T: T \in \mathcal{E}' \land \forall T'\in\mathcal{T}, T \gets T', T' \notin \mathcal{E}'\}$.
%    \EndFunction
%    \captionof{figure}{Adaptive parent selection.}
%\end{algorithmic}
%\end{center}

\tronly{
\noindent
\editremove{\textbf{Optimizations}}
\editinsert{%
\subsection{Optimizations}}
We implement some optimizations to help the system scale.
First, we use \emph{lazy updates} to the DAG, because the recursive definition for confidence may otherwise require a costly DAG traversal.
We maintain the current $d(T)$ value for each active vertex on the DAG, and update it only when a descendant vertex gets a chit.
Since the search path can be pruned at accepted vertices, the cost for an update is constant if the rejected vertices have limited number of descendants and the undecided region of
the DAG stays at constant size.
Second, the conflict set could be very large in practice, because a rogue client can generate a large volume of conflicting transactions.
Instead of keeping a container data structure for each conflict set, we create a mapping from each UTXO to the preferred transaction that stands as the representative for the entire conflict set.
This enables a node to quickly determine future conflicts, and the appropriate response to queries.
Finally, we speed up the query process by terminating early as soon as the $\alpha k$ threshold is met, without waiting for $k$ responses.
}{}

\editinsert{%
\subsection{Safety}
Lastly, the safety guarantees of Snowball can be mapped to those of Avalanche, which is a concrete instantiation of Snowball using a directed acyclic graph to amortize cost. 
We note that the structure of the Avalanche DAG itself does not correspond to votes, which is a subtle difference between other consensus protocols that make usage of a DAG. The DAG is merely a performance optimization, and is itself entirely orthogonal to the consensus process.
}

\editinsert{
\subsection{Liveness}
Avalanche introduces a DAG structure that entangles the fate of unrelated conflict sets, each of which is a single-decree instance.
This entanglement embodies a tension: attaching a virtuous transaction to undecided parents helps propel transactions towards a decision, while it puts transactions at risk of suffering liveness failures when parents turn out to be rogue.
We can resolve this tension and provide a liveness guarantee with the aid of two mechanisms.

First we adopt an adaptive parent selection strategy, where transactions are attached at the live edge of the DAG, and are retried with new parents closer to the genesis vertex. This procedure is guaranteed to terminate with uncontested, decided parents, ensuring that a transaction cannot suffer liveness failure due to contested, rogue transactions. 
A secondary mechanism ensures that virtuous transactions with decided ancestry will receive sufficient chits. Correct nodes examine the DAG for virtuous transactions that lack sufficient progeny and emit nop transactions to help increase their confidence.
With these two mechanisms in place, it is easy to see that, at worst, Avalanche will degenerate into separate instances of Snowball, and thus provide the same liveness guarantee for virtuous transactions.
}

\editinsert{%
\subsection{Communication Complexity}
Communication complexity for Avalanche is more subtle.
Let the DAG induced by Avalanche have an expected branching factor of $p$, corresponding to the width of the DAG, and determined by the parent selection algorithm.
Given the $\beta$ decision threshold, a transaction that has just reached the point of decision will have an associated progeny $\mathcal{Y}$.
Let $m$ be the expected depth of $\mathcal{Y}$.
If we were to let the Avalanche network make progress and then freeze the DAG at a depth $y$,
then it will have roughly $py$ vertices/transactions, of which $p(y - m)$ are decided in expectation.
Only $pm$ recent transactions would lack the progeny required for a decision.
For each node, each query requires $k$ samples, and therefore the total message cost per transaction is in expectation $(pky) / (p(y - m)) = ky/(y-m)$.
Since $m$ is a constant determined by the undecided region of the DAG as the system constantly makes progress, message complexity per node is $O(k)$, while the total complexity is $O(kn)$.
}

\section{Evaluation}\tronly{}{\vspace{-0.5em}}
\label{sec:evaluation}
\input{sections/evaluation}

\section{Related Work}\tronly{}{\vspace{-0.5em}}
\label{sec:related-work}
\input{sections/related-work}

\section{Conclusion}\tronly{}{\vspace{-0.5em}}
\label{sec:conclusions}
% This paper introduced a novel family of consensus protocols, coupled with the appropriate mathematical tools for analyzing them.
% Unlike classical or Nakamoto-style consensus, these protocols are highly efficient and robust. They scale well, achieve high throughput and low latency, work without precise membership knowledge, as well as degrade gracefully under catastrophic adversarial attacks, beyond current bounds.

% This work opens up new avenues of research. Chief among these is relaxing the synchrony assumption in this paper to accommodate asynchrony. We believe that the protocols and analysis techniques established in this paper will form the foundation for new venues of research.
% %and provide a strong safety guarantee,
% %though they achieve these properties by not guaranteeing liveness for conflicting transactions.
% %We have illustrated how they can be used to implement a Bitcoin-like payment system, that achieves 3400 tps in a geo-replicated setting.

This paper introduced a novel family of consensus protocols, coupled with the appropriate mathematical tools for analyzing them.
These protocols are highly efficient and robust, combining the best features of classical and Nakamoto-style consensus.
They scale well, achieve high throughput and quick finality, work without precise membership knowledge, and degrade gracefully under catastrophic adversarial attacks.

There is much work to do to improve this line of research. Chief among these is relaxing the synchrony assumption in this initial paper. We believe strong guarantees are possible even under asynchrony, but the details remain to be worked out. Another improvement would be to characterize the system's guarantees under an adversary whose powers are realistically limited, whereupon performance would improve even further.
Overall, we hope that the protocols and analysis techniques presented here add to the arsenal of the distributed system developers and provide a foundation for new lightweight and scalable mechanisms.

%\section*{Acknowledgments}\tronly{}{\vspace{-0.5em}}
%\label{sec:related-work}
%We are grateful to the many individuals who provided feedback on earlier drafts. 
% Bobby Kleinberg, Vitalik Buterin, ...

%latex here should be the name of your bibtex file minus '.bib'
\bibliographystyle{acm}
\bibliography{paper}
\newpage
\begin{appendices}
\section{Analysis}\tronly{}{\vspace{-0.5em}}
\label{sec:full-analysis}
In this appendix, we analyze Slush, Snowflake, Snowball, and Avalanche.
\tronly{}{For brevity, we leave details and proofs to an accompanying tech report~\cite{avalanche}.}

\paragraph{Network Model} We assume a synchronous communication network, where at each time step $t$,
a global scheduler chooses a single correct node uniformly at random.

\paragraph{Preliminaries} Let $c = |\mathcal{C}|$ and $b = |\mathcal{B}|$;
let $u \in \mathcal{C}$ be any correct node;
let $k \in \mathbb{N}_+$;
and let $\alpha \in \mathbb{R} = (1/2, 1]$.
We let $\mathtt{R}$ (``red'') and $\mathtt{B}$ (``blue'') represent two generic conflicting choices.
Without loss of generality, we focus our attention on counts of $\mathtt{R}$, i.e.\ the total number of nodes that prefer $\mathtt{R}$.

Each network query of $k$ peers corresponds to a sample without replacement out of a network of $n$ nodes, also referred to as a hypergeometric sample.
We let the random variable $\mathcal{H}_{\mathcal{N}, x}^k \rightarrow [0, k]$ denote the resulting counts of $\mathtt{R}$ (ranging from 0 to $k$), where $x$ is the total count of $\mathtt{R}$ in the population. The probability that the query achieves the required threshold of $\alpha k$ or more votes is given by:

\begin{equation}
\small
P(\mathcal{H}_{\mathcal{N}, x}^k \geq \alpha k) = \sum_{j = \alpha k}^{k} \frac{{x \choose j} {n - x \choose k - j}}{{n \choose k}}
\label{eq:hypergeometric}
\end{equation}

\tronly{
\paragraph{Tail Bound} We can reduce some of the complexity in Equation~\ref{eq:hypergeometric} by introducing a bound on the hypergeometric distribution induced by $\mathcal{H}^k_{\mathcal{C},x}$.
Let $p=x/c$ be the ratio of support for $\mathtt{R}$ in the population.
The expectation of $\mathcal{H}_{\mathcal{C}, x}^k$ is exactly $kp$.
Then, the probability that $\mathcal{H}_{\mathcal{C}, x}^k$ will deviate from the mean by more than some small constant $\psi$ is given by the Hoeffding tail bound \cite{hoeffding1963probability}, as follows,
\begin{equation}
\begin{split}
    P(\mathcal{H}_{\mathcal{C}, x}^k \leq (p-\psi)k) &\leq e^{-k\mathcal{D}(p-\psi, p)}
    % \leq e^{-2\psi^2k}
\end{split}
\end{equation}
where $\mathcal{D}(p-\psi, p)$ is the Kullback-Leibler divergence, measured as
\begin{equation}
\begin{split}
    \mathcal{D}(a, b) &= a \log \frac{a}{b} + (1 - a) \log \frac{1 - a}{1 - b}
\end{split}
\end{equation}}{}

\subsection{Analysis of Slush}\tronly{}{\vspace{-0.5em}}
%\label{sec:analysis}
Slush operates in a non-Byzantine setting; that is, $\mathcal{B} = \emptyset$ and thus $\mathcal{C} = \mathcal{N}$ and $c = n$.
In this section, we will show that
(R1) Slush converges to a state where all nodes agree on the same color in finite time almost surely (i.e.~with probability~1); 
(R2) provide a closed-form expression for speed towards convergence; and 
(R3) characterize the minimum number of steps per node required to reach agreement with probability $\ge 1-\varepsilon$.

\begin{figure}
    \small
\begin{algorithmic}[1]
    \State initialize $u.\codecolor \in \{\texttt{R},\texttt{B}\}$ for all $u\in\mathcal{N}$
    \For{$t = 1$ to $\phi$}
        \State $u \coloneqq \Call{sample}{\mathcal{N}, 1}$
        \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
        \State $P \coloneqq \texttt{[}v.\codecolor\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
        \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
            \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
                \State $u.\codecolor \coloneqq \codecolor'$
            \EndIf
        \EndFor
    \EndFor
\captionof{figure}{Slush run by a global scheduler.}\label{fig:slush_protocol_simulator}
\end{algorithmic}
\end{figure}

\tronly{
The procedural version of Slush in Figure~\ref{fig:slush-loop} made use of a parameter $m$, the number of rounds that a node executes Slush queries. To derive this parameter, we transform the protocol execution from a procedural and concurrent version to one carried out by a scheduler, shown in Figure~\ref{fig:slush_protocol_simulator}.
What we ultimately want to extract is the total number of rounds $\phi$ that the scheduler will need to execute in order to guarantee that the entire network is the same color, whp.
}{}

We analyze the system mainly using standard Markov chain techniques. Let $\mathbf{X} = \{\mathit{X_1}, \mathit{X_2}, \dots\, \mathit{X}_\phi\}$ be a discrete-time Markov chain.
Let $s_i$, $i \in [0, c]$, represent a state in this Markov chain, where the count of $\mathtt{R}$ is $i$ and the count of $\mathtt{B}$ is $c - i$.
Let $\mathit{M}$ be the transition probability matrix for the Markov chain.
Let $q(s_i) = \mathit{M}(s_i, s_{i-1})$, $p(s_i) = \mathit{M}(s_i, s_{i+1})$, and $r(s_i) = \mathit{M}(s_i, s_{i})$, where $p(s_0) = q(s_0) = p(s_c) = q(s_c) = 0$.
This is a birth-death process where the number of states is $c$ and the two endpoints, $s_0$ and $s_c$, are absorbing states, where all nodes are red and blue, respectively.

To extract concrete values, we apply the following probabilities to $q(s_i)$ and $p(s_i)$, which capture the probability of picking a red node and successfully sampling for blue, and vice versa.
\begin{equation*}
\begin{split}
&p(s_i) = \frac{c-i}{c}P(\mathcal{H}_{\mathcal{C}, i}^{k} \geq \alpha k)\\
&q(s_i) = \frac{i}{c}P(\mathcal{H}_{\mathcal{C}, c-i}^{k} \geq \alpha k)
\end{split}
\end{equation*}

\begin{lemma}[R1]
Slush reaches an absorbing state in finite time almost surely.
\end{lemma}
\begin{proof}
% This follows in a straightforward way from the construction of $\mathbf{X}$.
% More precisely,
Let $s_i$ be a starting state where $i \leq c - \alpha k$. All such states $s_i$ then have a non-zero probability of reaching absorbing state $s_0$ for all time-steps $t \geq i$. States where $i > c - \alpha$ have no possible threshold majority for $\mathtt{B}$, and all timesteps $t \leq i$ do not allow enough transitions to ever reach $s_0$. 
Under these conditions, the probability of absorption is strictly $> 0$, and therefore Slush converges in finite steps. 
\end{proof}

% First we note that this Markov chain is \emph{absorbing}, that is, every state in the chain can
% reach an absorbing state.  Moreover, because the chain has a finite number of states, the expected
% number of steps of reaching an absorbing state is finite.  Note also that the chain is symmetric,
% and one can think of this Markov chain as a generalization of the Gambler's Ruin Problem, for which
% we have closed form expressions for the expected number of steps.  Finally, note that our Markov
% chain should absorb faster than Gambler's Ruin because the ``pull'' toward an absorbing state increases
% with reduced distance to an absorbing state.

% R2 can be shown using the following lemma, which computes the expected number of steps to reach $s_0$, starting from $s_i$.

% \begin{comment}
% The next two lemmas establish recurrence relations that enable us to compute both the probability of absorbing in the all-red state and the expected time to reach that state.
% \tronly{
% \begin{lemma}
% \textbf{Probability of Deciding $\mathtt{R}$}. Let $\mathscr{A}(s_i)$ be the event of absorption at state $s_0$ starting from state $s_i$.
% Let
% \begin{equation}
%     \small
%     \mathscr{Y}_c(h) = \sum_{l = s_h}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
% \end{equation}
% Then,
% \begin{equation}
%     \small
%     P(\mathscr{A}(s_i)) = \mathscr{Y}_c(i) / \mathscr{Y}_c(0)
% \end{equation}
% % \begin{comment}
% % \begin{equation}
% % \begin{split}
% %     P(\mathscr{A}(s_i)) &= \mathscr{U}_n^{-1}\sum_{l = s_i}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
% % \end{split}
% % \end{equation}
% % where
% % \begin{equation}
% % \begin{split}
% %     \mathscr{U}_n = \sum_{l = s_0}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
% % \end{split}
% % \end{equation}
% % \end{comment}
% \label{lemma:absorption_theta_0}
% \end{lemma}
% \begin{proof}
% For $i \in [1, c)$, the absorbing probability satisfies the recurrence relation
% \begin{align}
%     P(\mathscr{A}(s_i)) =\ q(s_i)P(\mathscr{A}(s_{i-1})) &+ p(s_i)P(\mathscr{A}(s_{i+1})) \nonumber \\
%     &+ r(s_i)P(\mathscr{A}(s_i))
% \end{align}
% with boundary conditions $P(\mathscr{A}(s_0)) = 1$ and $P(\mathscr{A}(s_{c})) = 0$. Solving that recurrence relation yields the result.
% \end{proof}}{}
% \end{comment}

\tronly{
\begin{lemma}[R2]
\label{lemma:expected_absorption}
\textbf{Expected Rounds to Deciding $\mathtt{B}$.} Let $\mathit{U}_{s_i}$ be a random variable that expresses the number of steps to reach $s_0$ from state $s_i$.
Let
\begin{equation}
    \small
    \mathscr{Y}_c(h) = \sum_{l = s_h}^{s_{c-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}
\end{equation}
Then,
\begin{equation}
\small
\begin{split}
    \mathbb{E}[\mathit{U}_{s_i}] &= \mathbb{E}[\mathit{U}_{s_1}]\sum_{l=s_0}^{s_{i-1}}\prod_{j = s_1}^{l}\frac{q(j)}{p(j)}\\
    &- \sum_{l=s_0}^{s_{i-1}}\sum_{j=s_1}^{l}\frac{1}{p(j)}\prod_{v=j}^{l}\frac{q(v+1)}{p(v+1)}
\end{split}
\label{equation:expected_convergence}
\end{equation}
where
\begin{equation}
\small
\begin{split}
    \mathbb{E}[\mathit{U}_{s_1}] &= \mathscr{Y}_c(0)^{-1}\sum_{l=s_1}^{s_{c-1}}\sum_{j=s_1}^{l}\frac{1}{p(j)}\prod_{v=j}^{l}\frac{q(v+1)}{p(v+1)}
\end{split}
\end{equation}
\end{lemma}
\begin{proof}
The expected time to absorption in $s_0$ can be modeled using the following recurrence
equation:
\begin{align}
        \mathbb{E}[\mathit{U}_{s_i}] = q(s_i)\mathbb{E}[\mathit{U}_{s_{i-1}}] &+ p(s_i)\mathbb{E}[\mathit{U}_{s_{i+1}}]\nonumber \\
        &+ r(s_i)\mathbb{E}[\mathit{U}_{s_{i}}]
\end{align}
for $i \in [1, c-1]$, with boundary condition $\mathbb{E}[\mathit{U}_{s_0}] = 0$. 
Solving for the recurrence relation yields the result. 
\end{proof}
}{}

% \begin{lemma}
% Let $\phi \to \infty$. The Slush protocol enters an absorbing state at $\phi^{'}$, where $\phi^{'}$ is finite.
% \end{lemma}
% \begin{proof}
%     The chain $M$ is irreducible and aperiodic, and therefore converging. \tronly{More precisely, applying $p(s_i)$ and $q(s_i)$ computed above to Lemma~\ref{lemma:expected_absorption}, we see that $\phi^{'}$ is finite over all network configurations where count of $\mathtt{R}$ or count of $\mathtt{B}$ $\geq \alpha k$.}{}
% \end{proof}
% \tronly{

% \begin{lemma}
% \textbf{``Pull'' from State $s_i$.} Let $g(s_i) = q(s_i)/p(s_i)$ represent the ratio of the speed of absorption at $s_0$, starting at state $s_i$. Then, a lower-bound of $g(s_i)$ is
% \begin{equation}
% \small
% (c-i)^{\tau - 1}\ i^{1-\tau} \leq g(s_i)
% \end{equation}
% where $\tau = (2\alpha - 1) k$.
% \label{lemma:speed_absorption}
% \end{lemma}
% \begin{proof}
%     This follows directly from the lower bound of the binomial coefficient, i.e. $(z/k)^k \leq {n \choose k}$, for some $z \geq k \geq 1$, applied to the expression of $g(s_i)$.
% \end{proof}
% This inequality provides some insight into the convergence process.
% The bivalent state $s_{c/2}$, where the network is divided equally between the two colors, poses the worst case in terms of expected rounds to convergence.
% As the network diverges from the 50/50 state $s_{c/2}$, as measured by $i$, the probability of moving further in that direction increases at least polynomially in $i$.
% }{}

\begin{table}[h!]
	\centering
	\begin{tabular}{llllll}
		$\mathcal{C}$ Size   & 600    & 1200   & 2400   & 4800   & 9600    \\ \hline
		Exp. Conv. & 12.66 & 14.39 & 15.30 & 16.43 & 18.61
	\end{tabular}
    \caption{Expected number of per-node-iterations to convergence starting at worst-case (equal) $\mathcal{C}$ network split, in the case of $k = 10$, $\alpha=0.8$. Standard deviation for all samples is $\leq 2.5$. Computed via a Monte Carlo simulation.}
	\label{table:growth_worst_case}
\end{table}
Table~\ref{table:growth_worst_case} shows the latency (i.e. number of timesteps per-node) to reach the $\mathtt{B}$ deciding state starting from the metastable 50/50 split\tronly{, in other words, value $\mathbb{E}[\mathit{U}_{s_{c/2}}]/c$}{}. 
% \Jon{How was the table computed? Monte Carlo or otherwise numerically?}
As the network size $c$ doubles, expected convergence grows only linearly.

Finally, to demonstrate R3, we need to compute the first timestep of the Markov chain $M$ that yields a probability of $\leq \varepsilon$ for all transient states.
% Since the last $\alpha k - 1$ states (i.e. $s_{c-\alpha k}$, $\dots$, $s_{c}$) correspond to regions where nodes can no longer receive a sample of $\alpha k$ or more successes for the opposite color, the first of such states can serve as the desired target \tronly{state for computing the minimal time to $\varepsilon$-convergence.}{.}
This computation has no simple closed-form approximation, therefore we must iteratively raise the transition probability matrix $M$ until the first timestep $t$ that achieves probability $\leq \varepsilon$ for all states $s_i$ where $s_{\alpha k} < s_i < s_{c - \alpha k}$.
% \tronly{Figure~\ref{fig:expected_time_to_epsilon_convergence_slush} shows the expected time to reach $s_{c - \alpha k}$ whp, over various network sizes, with fixed $k = 10, \alpha = 0.8$.

% \begin{figure}
% \centering
% % file: gossip-chain/paper/kevinscripts/expected_time_to_epsilon_convergence_slush.py
% \tronly{\includegraphics[width=\linewidth]{figures/epsilon-convergence-large.pdf}}
%         {\includegraphics[width=\linewidth]{figures/epsilon-convergence.pdf}}
% \caption{Necessary parameter $\phi/c$ required in order to achieve $\varepsilon$-convergence in the Slush protocol, i.e.\ every node has reached the same color whp. We fix $k = 10$, $\alpha = 0.8$.}
% \label{fig:expected_time_to_epsilon_convergence_slush}
% \end{figure}
% }{}

\tronly{
The next two additional results provide some useful intuition to the underlying stochastic processes of Slush and subsequent protocols. 
The probability of a single successful query as a function of the total number of $\mathtt{R}$-supporting nodes in the network is shown in Figure~\ref{fig:varyingk}.
This probability decreases rapidly as the value of $k$ increases.
The network topples over faster if the value of $k$ is smaller due to the larger amount of random query perturbance.
Indeed, for Slush, $k=1$ is optimal.
Later we will see that for the Byzantine case, we will need a larger value for $k$.
}{}

\begin{figure}
\centering
% file: gossip-chain/paper/plots/trend-divergence.py
\tronly{\includegraphics[width=\linewidth]{figures/trend-diver-large.pdf}}
        {\includegraphics[width=\linewidth]{figures/trend-diver.pdf}}
\caption{Probability of successful samples vs.\ network splits.}
\label{fig:trend_divergence}
\end{figure}

Figure~\ref{fig:trend_divergence} shows the divergence of the probabilities of successful samples over various network splits. When the network is evenly split, the probabilities are equal for either color, but very rapidly diverge as the network becomes less balanced.

\begin{comment}
In this section, we have shown three results. First, we showed that Slush convergences to an absorbing state in finite time. Second, we provided a closed-form expression to calculate the expected time to convergence given any starting state, which provides useful information about the effectiveness of Slush to quickly terminate. Third, we discussed the iterative method, as well as some supporting numerical values, to calculate the minimum number of timesteps per node to guarantee absorption with probability $\geq 1 - \varepsilon$. 
\end{comment}

\tronly{
\begin{figure}[h!]
\centering
% file: gossip-chain/paper/kevinscripts/plots/varying_k_theta_0.py
\tronly{\includegraphics[width=\linewidth]{figures/varying-k-theta0-large.pdf}}
        {\includegraphics[width=\linewidth]{figures/varying-k-theta0.pdf}}
\caption{Probability of a single successful majority-red sample with varying $k$, $\alpha = 0.8$, and successes ranging from half the network to full-support.}
\label{fig:varyingk}
\end{figure}
}{}

\subsection{Safety Analysis of Snowflake}\tronly{}{\vspace{-0.5em}}
Snowflake, whose pseudocode is shown in Figure~\ref{fig:snowflake_simulator}, is similar to Slush, but has three key differences.\tronly{}{The scheduler version of Snowflake, omitted for brevity, is analogous to that for Slush, but has three key differences.} 
First, the sampled set of nodes includes Byzantine nodes.
Second, each node also keeps track of the total number of consecutive times it has sampled a majority of the same color.
We also introduce a function called $\mathcal{A}$, the adversarial strategy, that takes as parameters the entire space of correct nodes as well as the chosen node $u$ at time $t$, and as a side-effect, modifies the set of nodes $\mathcal{B}$ to some arbitrary configuration of colors.

The birth-death chain that models Snowflake, shown below, is nearly identical to that of Slush, with the added presence of Byzantine nodes. 
\begin{equation*}
% \small
\begin{split}
&p(s_i)=\frac{c-i}{c}P(\mathcal{H}_{\mathcal{N}, i}^{k} \geq \alpha k),\\
&q(s_i)=\frac{i}{c}P(\mathcal{H}_{\mathcal{N}, c-i+b}^{k} \geq \alpha k)
\end{split}
\end{equation*}
In Slush, the metastable state of the network was a 50/50 split at Markov state $s_{c/2}$. 
In Snowflake, Byzantine nodes have the ability to compensate for deviations from $s_{c/2}$ up to a threshold proportional to their size in the network.
Let the Markov chain state $s_{ps} > s_{c/2}$ be the state after which the probability of entering the $\mathtt{R}$-absorbing state becomes greater than the probability of entering the $\mathtt{B}$-absorbing state given the entire weight of the Byzantine nodes. 
Symmetrically, there exists a phase shift point for \texttt{B} as well. 
All states between these two phase-shift points are Byzantine controlled, whereas all other states are metastable.

In this section, we analyze the two conditions that ensure the safety of Snowflake. The first condition (C1) requires demonstrating that that there exists a point of no return $s_{ps + \Delta}$ after which
the maximum probability under full Byzantine activity of bringing the system back to the phase-shift point is less than $\varepsilon$.

The second condition (C2) requires the construction of a mechanism to ensure that a process can only commit to a color if the system is beyond the minimal point of no return, whp. 
As we show later, satisfying these two conditions is necessary and sufficient to guarantee safety of Snowflake whp.

\begin{figure}
\begin{algorithmic}[1]
    \State initialize $\forall u, u.\codecolor$,
    \State initialize $\forall u, u.\codecount \coloneqq 0$
\For{$t = 1$ to $\phi$}
    \State $u \coloneqq \Call{sample}{\mathcal{C}, 1}$
    % \State $u$.total++
    \State $\mathcal{A}(u, \mathcal{C})$
    \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
    \State $P \coloneqq \texttt{[}v.\codecolor\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
    \For{$\codecolor' \in \{\texttt{R}, \texttt{B}\}$}
    \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
        \If{$\codecolor' \neq u.\codecolor$}
            \State $u.\codecolor \coloneqq \codecolor'$, $u.\codecount \coloneqq 0$
        \Else
            \State $\texttt{++}u.\codecount$
        \EndIf
    \EndIf
    \EndFor
\EndFor
\captionof{figure}{Snowflake run by a global scheduler.}\label{fig:snowflake_simulator}
\end{algorithmic}
\end{figure}

\paragraph{Properties of Sampling Parameter $k$.} Before constructing and analyzing the two conditions that satisfy the safety of Snowflake, we introduce some important observations of the sample size $k$. 
The first observation of interest is that the hypergeometric distribution approximates the binomial distribution for large-enough network sizes.
If $x$ is a constant fraction of network size $n$, then:

% This is formalized in the following lemma:
% \begin{lemma}
% For $x$ as a fixed ratio of network support with increasing $n$, we have:
\begin{equation}
    \small
    \begin{split}
        \lim_{n \rightarrow \infty} P(\mathcal{H}_{\mathcal{N}, (xn)}^k \geq \alpha k) = \sum_{j=\alpha k}^k {k \choose j}x^k (1 - x)^{k - j}
    \end{split}
\end{equation}
% \end{lemma}
% \begin{comment}
% \begin{proof}
%     The proof follows in a straightforward way by applying the limit as $n \rightarrow \infty$ to the hypergeometric expression. For all $j = [\alpha k, k]$, we have:
%     \begin{equation}
%         \begin{split}
%             & \lim_{n \rightarrow \infty} {x \choose j}{n - x \choose k - j}/{n \choose k}\\
%             = & \lim_{n \rightarrow \infty} {k \choose j} \prod_{l = 1}^{j} \frac{x - j + l}{n - j + l} \prod_{m = 1}^{k - j} \frac{n - x - k + j + m}{n - k + m}\\
%             = & {k \choose j} (x/n)^{j} (1 - (x/n))^{k - j}
%         \end{split}
%     \end{equation}
%     and therefore, summed over all $j$, we have our approximation. 
% \end{proof}
% \end{comment}
\noindent This dictates that for large enough network sizes, the probability of success for any single sample becomes a function dependent only on the values of $k$ and $\alpha$, unaffected by further increases of network size. 
The result has consequences for the scalability of Snowflake and subsequent protocols. For sufficiently large and fixed $k$, the security of the system will be approximately equal for all network sizes $n > k$.
% Kevin: The next result bounds how fast the approximation occurs, but we don't widely make usage of this bound apart from providing intuition. Therefore, I think removing it completely is the sensible thing to do. 
% Following up, the next result bounds this approximation for small network sizes.
% \begin{lemma}
%     The probabilities of the hypergeometric experiment, given a fixed network sample parameter $k$ and $\alpha$ threshold, are independent of network size, up to a constant approximation factor $L$.
%     \label{lemma:lconvergence}
% \end{lemma}
% \tronly{
% \begin{proof}
%     To show this, we fix $k$ and $\alpha$ to two constants, and then compute the deviation of the probabilities of the hypergeometric as the network size (and successes $x$) increases by some factor $\gamma$. More precisely,
%     \begin{equation}
%         \small
%         \begin{split}
%             L = \lim_{\gamma \rightarrow \infty} \left( \sum_{j = \alpha k}^{k} \frac{{x\gamma \choose j} {c\gamma - x\gamma \choose k - j}}{{c\gamma \choose k}} \middle/ \sum_{j = \alpha k}^{k} \frac{{x \choose j} {c - x \choose k - j}}{{c \choose k}} \right)
%         \end{split}
%     \end{equation}
%     Rearranging, and using the smallest factor in the summation ($j = k$), we see that
%     \begin{equation}
%         \small
%         \begin{split}
%             L \geq L^{'} = \xi \lim_{\gamma \rightarrow \infty} \frac{{x\gamma \choose k}}{{c\gamma \choose k}}, \text{where}\ \xi = \frac{{c \choose k}}{{x \choose k}}
%         \end{split}
%     \end{equation}
%     For the lower bound of $x = c/2$ and the upper bound of $x = c$, we have that $2^{-k}\xi \leq L^{'} \leq \xi$.
%     % A perhaps cleaner, but less precise, way of exposing the relationship between $k$ and the network size is by using the Hoeffding tail bound, which shows that $P(\mathcal{H}_{\mathcal{C}, x}^k \leq ((x/c) - \psi)k) \leq e^{-2k((x/c) - \psi)^2}$ \cite{hoeffding1963probability}, where $\psi$ is a constant. As the network size increases, it has no effect on the probability distribution of the outcomes. Contrary to our more precise approximation of the bound above, the Hoeffding bound entirely removes 
% \end{proof}}{}
% \noindent The prior two lemmas indicate that Snowflake scales well with increasing network size. 
% In Figure \ref{fig:plot_k_needed_for_same_security_over_increasing_network} we show how $L$ converges over various network sizes. For networks of size $\geq 1000$, probabilities of various outcomes change minimally for the same $k$ and $\alpha$ parameters.
% \begin{figure}
% % file: gossip-chain/paper/kevinscripts/plot_k_needed_for_same_security_over_increasing_network.py
% \includegraphics[width=\linewidth]{./plot_k_needed_for_same_security_over_increasing_network.pdf}
% \caption{Ratios of the probabilities of successes over various network sizes, with fixed $k = 10$, $\alpha = 0.8$, and $x = 0.8c$. \texttt{RVR: the labels are too small, and we don't need a title above the graph.}}
% \label{fig:plot_k_needed_for_same_security_over_increasing_network}
% \end{figure}

We now examine how the phase shift point behaves with increasing $k$. %At the maximum allowable $k = n$, the phase shift point $s_{ps}^{*}$, reaches its maximal value.
Whereas $k=1$ is optimal for Slush (i.e. the network topples the fastest when $k = 1$), we see that Snowflake's safety is maximally displaced from its ideal position for $k=1$, reducing the number of feasible solutions.
Luckily, small increases in $k$ have an exponential effect on $s_{ps}$, creating a larger set of feasible solutions. More formally: 
\begin{lemma}
$s_{ps}$ approaches $s_{n/2}$ exponentially fast as $k$ approaches $n$, and at $k = n$, $s_{ps} = s_{n/2}$.
\end{lemma}
\begin{proof}
At $s_{ps}$, the ratio $p(s_i)/q(s_i)$ is equal to 1, by definition. Substituting the solution $i = c/2 + b/2$ into this equation yields the result at $k=n$.
Computing the differential in $s_{ps}$ between any chosen $k$ and $k+1$ yields the exponential trend.
\end{proof}

% \noindent Figure~\ref{fig:plt_various_k_phase_shift} shows the exponential relationship between decreasing $k$ and $s_{ps}$. The graph shows that small-but-not-tiny values of $k$ lead to large feasible regions.

% \begin{figure}[h!]
%     \centering
%     % file: gossip-chain/paper/kevinscripts/plots/plot_various_k_phase_shift.py
% \tronly{\includegraphics[width=\linewidth]{figures/k-phase-shift-large.pdf}}
%         {\includegraphics[width=\linewidth]{figures/k-phase-shift.pdf}}
%     \caption{Phase shift state of Snowflake/Snowball vs. $k$ values (x-axis), with $n = 2000$, $\alpha = 0.8$ and $b = 20\%$ of $n$. The phase-shift is indexed starting from $c/2$. \Jon{I really didn't get the story here. It seems like cool graphical representations that would really help me understand the underlying arguments, but I missed something.}}
%     \label{fig:plt_various_k_phase_shift}
% \end{figure}

% To analyze the probability of the Byzantine nodes triggering safety violation, we use the same state space $\forall i\,\,\{s_i\}$ as in Slush, with the difference that the transition probabilities are now dependent on the strategy chosen by the Byzantine nodes, complicating analysis.
% We sidestep this difficulty by conservatively focusing on the states that are irreversible under \emph{any} Byzantine strategy.

% {
% \color{red}
% Let $\Delta$ represent the network bias towards red, represented by state $s_{c/2 + \Delta}$.
% Let $s_{ps} \geq s_{c/2}$ represent the phase-shift state in the system, as defined previously.
% We are interested in determining network parameters $k$ and $\alpha$ for an anticipated maximum number of Byzantine nodes $b$. Then, $\exists \Delta$ where $s_{c/2 + \Delta} > s_{ps}$ s.t.\ the probability of the system reverting back to $s_{ps}$ is less than or equal to $\varepsilon$ for some upper bound $\phi$. Figure~\ref{fig:states_feasible_solutions} summarizes the region of interest,
% where the desired $\Delta$ is the \emph{point of no return} after which the network is not likely to revert whp. We refer to states that meet this condition with regards to $s_{ps}$ as \emph{irreversible}.
% }

\paragraph{Satisfiability of Safety Criteria}
Let $\varepsilon$ be a system security parameter of value $\leq 2^{-32}$, typically referred to as \emph{negligible} probability.

% $\textsc{isAccepted}$ be the function which returns true if $u$ samples the same color successfully for at least $\beta$ consecutive times.
% A crucial property of $\textsc{isAccepted}$ is that it returns true 
% We are interested in the set of outcomes where, at time $\phi$, there exists at least two distinct correct nodes $u$ and $v$ for which the \textsc{isAccepted} predicate returns true for different colors. \tronly{Formally, we are interested in the following game:
% \begin{equation}
%     \begin{cases}
%         \textrm{win} & \textbf{if}\ t = \phi \land \exists u, v, C, C': u \ne v \land C \ne C' \Rightarrow \\
%                      & u.\textsc{isAccepted}(C) \land v.\textsc{isAccepted}(C') \\
%         \textrm{lose} &  \textbf{otherwise}
%     \end{cases}
% \end{equation}}{}
We now formalize and analyze the safety condition C1. We refer to all states $s_i$ that satisfy condition C1 as \emph{states of feasible solutions}. 
For the system to guarantee safety, there must exist some state $s_{c/2 + \Delta} > s_{c/2 + b} > s_{ps}$ that implies $\varepsilon$-irreversibility in the birth-death chain.
In other words, once the system reaches this state, then it reverts back with only negligible probability, under \emph{any} Byzantine strategy. Figure~\ref{fig:states_feasible_solutions} summarizes the region of interest.
Our goal is to determine values for tunable system parameters to achieve this condition, given target network size $n$, maximum Byzantine component $b$, and security parameter $\varepsilon$, fixed by the system designer.

\begin{figure}[h]
    \begin{center}
\input{figures/analysis1}
\caption{Representation of the space of feasible solutions. Outside the phase-shift point, the network builds a bias towards one end. At the $\Delta$ point, this bias is $\varepsilon$-irreversible.
% \Jon{I really didn't get the story here. It seems like cool graphical representations that would really help me understand the underlying arguments, but I missed something.}
}
\label{fig:states_feasible_solutions}
    \end{center}
\end{figure}

\noindent To ensure C1, we must find $\Delta$ such that the following condition holds:
\begin{equation}
\begin{split}
    \exists \Delta, \text{where}\ s_{c/2 + \Delta} &> s_{ps},\ s.t.\ \forall t \leq \phi:\\
    \varepsilon &\geq (V_c[s_{c/2 + \Delta}, s_{ps}])^t
\end{split}
\end{equation}
where $V_c$ is the $(c+1)^2$ probability transition matrix for the system.
In probability theory terminology, the statement above computes the $t$-step hitting probability, where the starting state is $s_{c/2 + \Delta}$ and the hitting state is $s_{ps}$.
Satisfying this first constraint, that is, moving the ensemble of nodes to a decision, is not sufficient. The nodes have to be able to assure \emph{themselves} that they can safely decide, which means they must ensure condition C2.
While there are multiple ways of constructing the predicate that ensures this condition, the one we adopt in Snowflake involves $\beta$ consecutive successful queries, such that $P(\textsc{isAccepted} \,|\, s_i \leq s_{c/2 + \Delta}) \leq \varepsilon$. 
To achieve this threshold probability of failure for C2, we solve for the smallest $\beta$ that meets inequality
$P(\mathcal{G}(\mathtt{p}, \phi/c) \geq \beta) \leq \varepsilon$
where $\mathtt{p} = P(\mathcal{H}_{\mathcal{N}, i}^k \geq \alpha k)$ and $\mathcal{G}(\mathtt{p}, \phi/c)$ is a random variable which counts the largest number of consecutive same-color successes for a run over $\phi/c$ trials.
Solving for the minimum $i$ yields $\beta$.

The task of the system designer, finally, is to derive values for $\alpha$ and $\beta$, given desired $n, b, k, \varepsilon$.
Choice of $\alpha$ immediately follows from $b$ and is $(n - b)/ n$, the maximum allowable size to tolerate $\leq b$ Byzantine nodes.
Solving for $\Delta$ with a closed form expression would be ideal but unfortunately is difficult, hence we employ a numerical, iterative search.
Since $k$ and $\beta$ are mutually dependent, a system designer will typically fix one and search for the other.
Because $\beta$ is a tunable parameter whose consideration is primarily latency vs.\ security, a designer can fix it at a desired value and search for a corresponding $k$.
Since low $k$ are desirable, we perform this by starting at $k=1$, and successively evaluating larger values of $k$ until a feasible solution is found.
Depending on $b$ and the chosen $\varepsilon$, such a solution may not exist. If so, this will be apparent during system design. 
% \Jon{I found myself afloat in a sea of single-letter variable names, and was wishing for a quick-reference key to decode them. (Or even better, for some to be eliminated. :V)}

We finalize our analysis by formally composing C1 and C2: 
\begin{theorem}
    If C1 and C2 are satisfied under appropriately chosen system parameters, then the probability that two nodes $u$ and $v$ decide on $\mathtt{R}$ and $\mathtt{B}$ respectively is strictly $< \varepsilon$ over all timesteps $0 \leq t \leq \phi$. 
\end{theorem}
\begin{proof}
    The proof follows in a straightforward manner from the core guarantees that C1 and C2 provide. Without loss of generality, suppose a single node $u$ decides $\mathtt{R}$ at time $t < \phi$, and suppose that the network is at state $s_j$ at the time of decision. By construction, $\textsc{isAccepted}$ will return true with probability no greater than $\varepsilon$ for all network states $s_i < s_{c/2 + \Delta}$. Therefore, since $u$ decided, it must be the case that $s_j > s_{c/2 + \Delta}$, whp. Lastly, since the system will only revert back to a state with a majority of $\mathtt{B}$ nodes with negligible probability once it is past $s_{c/2 + \Delta}$, the probability that a node $v$ decides on $\mathtt{B}$ is strictly $< \varepsilon$.
\end{proof}

For\ \ illustration purposes, we examine a network with $n = 2000,\ \alpha = 0.8$, $\varepsilon \leq 2^{-32}$, and throughput of 10000tps.
If we allow this system to run for \textasciitilde{}4,000 years (which corresponds to $\phi = 10^{15}$), we choose $\beta$ such that the probability of a node committing
at state $s_{c/2+\Delta - 1}$ is $< \varepsilon$.
A safety failure then occurs when both a node commits $\mathtt{R}$ and the system reverts to $\mathtt{B}$. Figure~\ref{figure:failure_theta_0} shows the probability of failure and the corresponding $\beta$ for different choices of $k$.

\begin{figure}[h!]
\tronly{\includegraphics[width=\linewidth]{figures/consecutive-beta-large.pdf}}
        {\includegraphics[width=\linewidth]{figures/consecutive-beta.pdf}}
    \caption{$\beta$ and failure probability $p_f$ vs. $k$ values (x-axis), with $n = 2000$, $\alpha = 0.8$.}
\label{figure:failure_theta_0}
\end{figure}

\begin{comment}
In this section, we have shown that if the network parameters chosen by the system designer satisfy both conditions C1 and C2, then probability of a safety failure in Snowflake is negligible. 
Finally, there exists one more consequence to the Snowflake protocol.
Namely,
\end{comment}
Note that a larger choice of the $\alpha$ parameter dictates a larger space of feasible solutions, which implies a larger tolerance of Byzantine presence. 
If the system designer chooses to allow a larger presence of Byzantine nodes, she may achieve this goal at the cost of liveness. 
%XXX expand on this counterintuitive idea at some point, it implies that we can put up with very large byzantine coalitions
%XXX how big can $b$ be?
%XXX KEVIN: There are actually many cool things I can be saying additionally here, but what is written up to now is sufficient for tech report. I will soon-enough add up some interesting insights about speed of convergence once the system actually enters s_{c/2 + \Delta}.

\subsection{Safety Analysis of Snowball}\tronly{}{\vspace{-0.5em}}
In this section, we will demonstrate that Snowball provides strictly better
safety than Snowflake.
The key difference in protocol mechanics between Snowflake and Snowball is that correct nodes now monotonically increase their confidence in $\mathtt{R}$ and $\mathtt{B}$.
This limits the effect of random perturbations, as well as the impact of the Byzantine adversary.

To intuitively understand the behavioral differences between Snowball and its predecessor, we illustrate with an example, accompanied by some concrete numerical values.
\begin{comment}
The example will also provide intuition on how Snowball admits a safety guarantee strictly stronger than that of Snowflake. The goal of this section is to formally prove this guarantee.
\end{comment}
Suppose a network of size $n$ where $b = (1/5)n$, $\alpha = (4/5)k$ (i.e. the maximum allowable threshold to tolerate the $(1/5)n$ presence of Byzantine nodes), and where $k$ is chosen large enough to admit a feasible solution. 
Suppose that the network swings to state $s_{c/2 + \Delta}$, such that correct nodes have a non-negligible probability of deciding on $\mathtt{R}$, and suppose, for the sake of simplicity, that $\Delta$ is chosen as the highest possible value, specifically $c/2 + \Delta + b = \alpha n = (4/5)n$, the upper-threshold set up by the system designer.

In this configuration, at least $(4/5)n - b = (3/5)n$ (correct) nodes prefer $\mathtt{R}$. 
These nodes perceive network support for $\mathtt{R}$ of at most $(4/5)n$ if the Byzantine nodes choose to also vote for $\mathtt{R}$, and at least $(3/5)n$ if the Byzantine nodes choose to withhold support for $\mathtt{R}$. 
Conversely, the rest of the $(1/5)n$ (correct) nodes that may prefer $\mathtt{B}$ perceive network support for $\mathtt{B}$ of at least $(1/5)n$ and at most $(2/5)n$, depending on the Byzantine strategy. 
Regardless of Byzantine strategy, however, all correct nodes that prefer $\mathtt{R}$ will have an expected growth of the $\mathtt{R}$ confidence to be strictly greater than the $\mathtt{B}$ confidence for the nodes that prefer $\mathtt{B}$. 
This is because network support for $\mathtt{R}$ is at least $(3/5)n$, which is greater than the maximum network support for $\mathtt{B}$ of $(2/5)n$. 
Therefore, once the network swings to the minimum required state, the expected growth of the $\mathtt{R}$-confidence will be greater than that of $\mathtt{B}$-confidence, for all correct nodes and for all Byzantine strategies.

In the previous section, we provided the two conditions necessary for guaranteeing the safety of Snowflake. The second condition, C2, remains identical in Snowball, and therefore does not need to be modified and re-analyzed. Conversely, condition C1 (i.e. showing that the network is irreversible whp) needs to be re-analyzed. In this section, using the same intuition provided in the example above, we formally show how Snowball achieves irreversibility under strictly higher probability than in Snowflake.

\tronly{
\paragraph{Security of Snowball.}
Figure~\ref{fig:snowball_simulator} illustrates the scheduler-based version of Snowball.
In Snowball, a node $u$ \textit{prefers} $\mathtt{R}$ if $u.d[\mathtt{R}] > u.d[\mathtt{B}]$, and vice versa. At the start of the protocol, these preferences are initialized to $\langle 0, 0 \rangle$. 
This is in contrast to Snowflake, where nodes prefer a color based only on the latest successful sample.
We can model the protocol with a Markov chain similar to that of Snowflake, and derive the parameters and feasibility for the protocol.

\begin{figure}[h!]
\begin{algorithmic}[1]
\State initialize $\forall u, u.\codecolor$
\State initialize $\forall u, u.\textrm{lastcol}$
\State initialize $\forall u, u.\codecount \coloneqq 0$
\State initialize $\forall u, \forall \codecolor' \in \{\texttt{R}, \texttt{B}\}, u.d[\codecolor'] \coloneqq 0$
\For{$t = 1$ to $\phi$}
    \State $u \coloneqq \Call{sample}{\mathcal{C}, 1}$
    \State $\mathcal{A}(u, \mathcal{C})$
    \State $\mathcal{K} \coloneqq \Call{sample}{\mathcal{N}\textbackslash u, k}$
    \State $P \coloneqq \texttt{[}v.\codecolor\quad\textbf{for}\ v \in \mathcal{K}\texttt{]}$
    \For{$c' \in \{\texttt{R}, \texttt{B}\}$}
    \If{$P.\Call{count}{\codecolor'} \ge \alpha\cdot k$}
        \State $u.d[\codecolor']\texttt{++}$
        \If {$u.d[\codecolor'] > u.d[\codecolor]$}
            \State $u.\codecolor \coloneqq \codecolor'$
        \EndIf
        \If{$\codecolor' \neq u.\textrm{lastcol}$}
            \State $u.\textrm{lastcol} \coloneqq \codecolor'$, $u.\codecount \coloneqq 0$
        \Else
            \State $\texttt{++}u.\codecount$
        \EndIf
    \EndIf
    \EndFor
\EndFor
\captionof{figure}{Snowball run by a global scheduler.}
\label{fig:snowball_simulator}
\end{algorithmic}
\end{figure}

We fix a $\Delta$ such that, under the Markovian construction of Snowflake, if the system reaches $s_{c/2 + \Delta}$, then whp it does not revert. We show that this choice of $\Delta$ provides an even stronger guarantee once the network switches protocol to Snowball.
Without loss of generality, we cluster the correct nodes into two groups, and represent each group through two nodes $u$ and $v$, where $u$ represents the set of correct nodes that prefer red and $v$ those that prefer blue. 
Lastly, let the state of the system be initialized at $s_{c/2 + \Delta}$.

% With non-negligible probability, $u$ will decide on $\mathtt{R}$, since the network is biased enough to trigger the decision predicate. 
The best possible confidence-configuration that the Byzantine nodes can attempt to force correct nodes into is where all $\mathtt{R}$-preferring nodes, represented by $u$, are forced to maintain nearly equal confidence between the two colors, and where all of the $\mathtt{B}$-preferring nodes, represented by $v$, gain as much confidence as possible for~$\mathtt{B}$.

Let $u.d[\mathtt{R}] = u.d[\mathtt{B}] + 1$, corresponding to the minimal viable color difference for the red-preferring nodes, and let $v.d[\mathtt{B}] - v.d[\mathtt{R}] = \kappa$ be the difference for the blue-preferring nodes. While $\kappa \geq 0$, then the expectation of preference growths at time $t$ are:
\begin{equation}
    {\scriptsize
\begin{split}
    \mathbb{E}[{u.d}^t[\mathtt{R}]] &= \mathbb{E}[{u.d}^{t-1}[\mathtt{R}]] + \left(\frac{1}{2}+\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 + \Delta + b}^k  \geq \alpha k)\\
    \mathbb{E}[{u.d}^t[\mathtt{B}]] &= \mathbb{E}[{u.d}^{t-1}[\mathtt{B}]] + \left(\frac{1}{2}+\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 - \Delta + b}^k  \geq \alpha k)\\
    \mathbb{E}[{v.d}^t[\mathtt{R}]] &= \mathbb{E}[{v.d}^{t-1}[\mathtt{R}]] + \left(\frac{1}{2}-\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 + \Delta}^k  \geq \alpha k)\\
    \mathbb{E}[{v.d}^t[\mathtt{B}]] &= \mathbb{E}[{v.d}^{t-1}[\mathtt{B}]] + \left(\frac{1}{2}-\frac{\Delta}{c}\right)P(\mathcal{H}_{\mathcal{N}, c/2 - \Delta + b}^k  \geq \alpha k)\\
\end{split}
    }
    \label{equation:expectation_snowball}
\end{equation}
Note that we are implicitly using an indicator function when adding up the probability of success to the expected confidence growth, where the expected value of the indicator value is exactly the probability of success of a sample. 
Substituting for Hoeffding's tail bound, we get the rate at which $\kappa$ decreases per time-step to be
\begin{equation}
    % p = \cfrac{c/2 + \Delta + b}{c+b}
    \begin{split}
    \small
        r = \biggl(\left(\frac{1}{2}-\frac{\Delta}{c}\right) &e^{-k\alpha \log \left(\alpha/\frac{c/2 - \Delta + b}{c+b}\right)}\\
    &e^{-k(1-\alpha) \log \left(\left(1-\alpha\right)/\left(1-\frac{c/2 - \Delta + b}{c+b}\right)\right)}\biggr)\\
        - \biggl(\left(\frac{1}{2}-\frac{\Delta}{c}\right)&e^{-k\alpha \log \left(\alpha/\frac{c/2 + \Delta + b}{c+b}\right)}\\
    &e^{-k(1-\alpha) \log \left(\left(1-\alpha\right)/\left(1-\frac{c/2 + \Delta + b}{c+b}\right)\right)}\biggr)\\
    \end{split}
    \label{equation:rate_decrease}
\end{equation}

Upon reaching $\kappa = -1$, the blue-preferring nodes will have flipped to red.
From then on, all correct nodes are red, and the confidence of red grows significantly faster than that of blue.
We now show that the rate of growth of $\mathtt{R}$ will always be larger than the rate of growth of $\mathtt{B}$, over all correct nodes.

Letting $X({u.d}^t[\mathtt{R}])$ be a random variable that outputs the confidence of red for $u$ at timestep $t$, and letting $\bar X({u.d}^t[\mathtt{R}]) = X({u.d}^1[\mathtt{R}]) + \dots + X({u.d}^t[\mathtt{R}])$, we apply Hoeffding's concentration inequality to get that $P({\bar X({u.d}^t[\mathtt{R}])} - \mathbb{E}[\bar X({u.d}^t[\mathtt{R}])] \geq l) \leq e^{-2l^2t}$.
At time $t$, the expected confidence of red for $u$ is
\begin{equation}
    {u.d}^0[\mathtt{R}] + t\left(\frac{1}{2} + \frac{\Delta}{c}\right)\left(\frac{\frac{c}{2} + \Delta + b}{c+b}\right)
\end{equation}
And conversely the expected confidence of blue for $u$ is
\begin{equation}
    {u.d}^0[\mathtt{B}] + t\left(\frac{1}{2} + \frac{\Delta}{c}\right)\left(\frac{\frac{c}{2} - \Delta + b}{c+b}\right)
\end{equation}
Then, the probability that the expected value of red confidence is close to the expected value of the blue confidence is
\begin{equation}
    e^{-2t^3(1/2 + \Delta/c)^2(2\Delta/(c+b))^2}
    \label{equation:concentration_mean}
\end{equation}
Solving for $t$ that leads to such probability being negligible leads to:
\begin{equation}
    t = \left(\cfrac{\log(\varepsilon)}{-2\cdot(1/2 + \Delta/c)^2(2\Delta/(c+b))^2}\right)^{1/3}
    \label{equation:minimal_time}
\end{equation}
Therefore, at only a small number of timesteps $t$, $u$ has a red-confidence centered around its mean, with probability $\varepsilon$ close to the mean of the blue-confidence. An identical analysis follows for $v$.
In other words, after a small number of time-steps, the red-confidence of $u$ grows large enough such that the probability of this value being close to that of the blue-confidence is negligible. Further timesteps amplify this distance and further decrease the probability of the two confidences being close. 
}

\tronly{}{The details are provided in an accompanying technical report~\cite{avalanche}. The central result is as follows:}
We conclude our result with the following Theorem:
\begin{theorem}
Over all viable network parameters $n,\ b$, and for all appropriately chosen system parameters $k$ and $\alpha$, the probability of violating safety in Snowball is strictly less than the probability of violating safety in Snowflake.
\end{theorem}
\begin{proof}
%    The proof is straightforward.
Under the construction of Snowball in comparison to Snowflake, we see that safety criterion C2 remains the same. In other words, the consecutive-successes decision predicate has the same guarantees. 
On the other hand, the probabilistic guarantees of C1 change, meaning that the reversibility probabilities of the system are different. 
However, as we determined in Equation~\ref{equation:minimal_time}, over a few timesteps
$u$'s confidence difference between $\mathtt{R}$ and $\mathtt{B}$ grows large enough to guarantee that this confidence difference will revert back with only negligible probability.
As time progresses, the probability of being close decreases poly-exponentially, as shown in Equation \ref{equation:concentration_mean}. The same results follow for $v$, but inversely, meaning that the means get closer to each other rather than deviate. This continues until $v$ flips color. 
\end{proof}

\tronly{Snowball has strictly stronger security guarantees than Snowflake, which implies that appropriately chosen parameters chosen for Snowflake automatically apply for Snowball. Using the same techniques as before, the system designer chooses appropriate $k$ and $\beta$ values that ensure the desired $\varepsilon$ safety guarantees.}{}

% ------------------------------------------------------
% OLD STUFF ABOUT THETA 1 AND BYZANTINE OPTIMALITY ADDED
% TO FILE oldtheta1.tex, see line below. DO NOT DELETE!!
% \input{oldtheta1.tex}
% ------------------------------------------------------

\subsection{Safety Analysis of Avalanche}\tronly{}{\vspace{-0.5em}}
The key difference between Avalanche and Snowball is that in Avalanche, queries on the DAG on transaction $T_i$ are used to implicitly query the entire ancestry of $T_i$.
In particular, a transaction $T_i$ is preferred by $u$ \textit{if and only if} all ancestors are also preferred.
Suppose that $T_i$ and $T_j$ are in the same conflict set.
We can now infer two things.
First, we can consider the entire ancestry of $T_i$ and $T_j$ as a single decision instance of Snowball, where the ancestry of $T_i$ can be considered to be the $\mathtt{R}$ decision, and the ancestry of $T_j$ can be considered to be the $\mathtt{B}$ decision.
Second, since $T_i$ must be preferred if a child of $T_i$ is to be preferred, then we can collapse the progeny of $T_i$ into a single urn which repeatedly adds confidence to $\mathtt{R}$ whenever a child of $T_i$ gets a chit.
Consequently, Avalanche maps to an instance of Snowball, with the previously shown safety properties.

We note, however, since a decision on a virtuous transaction is dependent on its parents, Avalanche's liveness guarantees do not mirror that of Snowball. We address this in the next two sub-sections.

\subsection{Safe Early Commitment}\tronly{}{\vspace{-0.5em}}
\tronly{
As we reasoned previously, each conflict set in Avalanche can be viewed as an instance of Snowball, where each progeny instance iteratively votes for the entire path of the ancestry.
This feature provides various benefits; however, it also can lead to some virtuous transactions that depend on a rogue transaction to suffer the fate of the latter.
%has the drawback that it can entangle the fate of some unfortunate virtuous transactions with the fate of rogue ones.
In particular, rogue transactions can interject in-between virtuous transactions and reduce the ability of the virtuous transactions to ever reach the required $\textsc{isAccepted}$ predicate.
}{}
As a thought experiment, suppose that a transaction $T_i$ names a set of parent transactions that are all decided, as per local view.
If $T_i$ is sampled over a large enough set of successful queries without discovering any conflicts, then, since by assumption the entire ancestry of $T_i$ is decided, it must imply that at least $c/2 + \Delta$ correct nodes also vote for $T_i$, achieving irreversibility.

To then statistically measure the assuredness that $T_i$ has been accepted by a large percentage of correct nodes without any conflicts, we make use of a one-way birth process, where a birth occurs when a new correct node discovers the conflict of $T_i$. \tronly{Necessarily, deaths cannot exist in this model, because a conflicting transaction cannot be unseen once a correct node discovers it. Let $t = 0$ be the time when $T_j$, which conflicts with $T_i$, is introduced to a single correct node $u$.
Let $s_x$, for $x = 1$ to $c$, be the state where the number of correct nodes that know about $T_j$ is $x$, and let $p(s_x)$ be the probability of birth at state $s_x$. Then, we have:
\begin{equation}
    p(s_x) = \frac{c - x}{c} \left(1 - \frac{{n - x \choose k}}{{n \choose k}}\right)
\end{equation}
Solving for the expected time to reach the final birth state provides a lower bound to the $\beta_1$ parameter in the $\textsc{isAccepted}$ fast-decision branch.}{We leave details to the tech report.} The table below shows an example of the analysis for $n = 2000$, $\alpha = 0.8$, and various $k$, where $\varepsilon \ll 10^{-9}$, and where $\beta$ is the minimum required size of $d(T_i)$.
\tronly{}{\vspace{-2ex}}
\begin{table}[h!]
    \small
	\centering
	\begin{tabular}{llllll}
		$k$   & 10 & 20 & 30 & 40 \\ \hline
		$\beta$ & 10.87625 & 10.50125 & 10.37625 & 10.25125
	\end{tabular}
	\label{table:fast-path-beta}
\end{table}
\tronly{}{\vspace{-2ex}}
% \begin{figure}
%     \includegraphics[width=\linewidth]{figures/fast-path-beta.pdf}
%     \captionof{figure}{An example of $\beta$ solutions for different $k$, with $n = 2000, \alpha = 0.8$.}
%     \label{fig:fast-path-beta}
% \end{figure}
\noindent Overall, a very small number of iterations are sufficient for the safe early commitment predicate.

\subsection{Churn and View Updates}\tronly{}{\vspace{-0.5em}}
\tronly{Any realistic system needs to accommodate the departure and arrival of nodes.}{}
Up to now, we simplified our analysis by assuming a precise knowledge of network membership.
We now demonstrate that Avalanche nodes \tronly{can admit a well-characterized amount of churn,
by showing how to pick parameters such that Avalanche nodes}{} can differ in their view of the network and still safely make decisions.

Consider a network whose operation is divided into epochs of length $\tau$, and a view update from epoch $t$ to $t+1$ during which $\gamma$ nodes join the network and $\bar \gamma$ nodes depart.
Under our static construction, the state space $s$ of the network had a key parameter $\Delta^{t}$ at time $t$, induced by $c^{t}, b^{t}, n^{t}$ and the chosen security parameters.
This can, at worst, impact the network by adding $\gamma$ nodes of color $\mathtt{B}$, and remove $\bar \gamma$ nodes of color $\mathtt{R}$.
At time $t+1$, $n^{t+1} = n^{t} + \gamma - {\bar \gamma}$, while $b^{t+1}$ and $c^{t+1}$ will be modified by an amount $\leq \gamma - {\bar \gamma}$, and thus induce a new $\Delta^{t+1}$ for the chosen security parameters.
This new $\Delta^{t+1}$ has to be chosen such that $P(s_{c^{t+1}/2 + \Delta^{t+1} - \gamma} \rightarrow s_{ps}) \leq \varepsilon$, to ensure that the system will converge under the previous pessimal assumptions. The system designer can easily do this by picking an upper bound on $\gamma, \bar \gamma$.

The final step in assuring the correctness of a view change is to account for a mix of nodes that straddle the $\tau$ boundary. We would like the network to avoid an unsafe state no matter which nodes are using the old and the new views.
The easiest way to do this is to determine $\Delta^t$ and $\Delta^{t+1}$ for desired bounds on $\gamma, \bar \gamma$, and then to use the conservative value $\Delta^{t+1}$ during epoch $t$. In essence, this ensures that no commitments are made in state space $s^t$ unless they conservatively fulfill the safety criteria in state space $s^{t+1}$. As a result, there is no possibility of a node deciding red at time $t$, the network going through an epoch change and finding itself to the left of the new point of no return $\Delta^{t+1}$.

This approach trades off some of the feasibility space, to add the ability to accommodate $\gamma, \bar \gamma$ node churn per epoch.
\tronly{
Overall, if $\tau$ is in excess of the time required for a decision (on the order of minutes to hours), and nodes are loosely synchronized,
they can add or drop up to $\gamma, \bar \gamma$ nodes in each epoch using the conservative process described above. }{}
% Due to space reasons,
We leave the precise method of determining the next view to a subsequent paper, and instead
rely on a membership oracle that acts as a sequencer and $\gamma$-rate-limiter, using technologies like Fireflies~\cite{JohansenRVJ15}.
% and Ethereum~\cite{wood2014ethereum}.
\end{appendices}
\end{document}