### 2020-11-07 avalanche_fr.tex
S5 - début DAG, création de avalanche_fr.tex

\subsection{Avalanche: Adding a DAG}%\tronly{}{\vspace{-0.5em}}

%O Avalanche consists of multiple single-decree Snowball instances instantiated as a multi-decree protocol that
%O maintains a dynamic, append-only directed acyclic graph (DAG) of all known transactions.
Avalanche consiste en plusieurs instance Snowball à décret unique instanciées en tant que protocole multi-décrets qui maintient un graphe orienté acyclique (\emph{Directed Acyclic Graph},  DAG) dynamique en ajout seul de toutes les transactions connues.
%O The DAG has a single sink that is the \emph{genesis vertex}.
Le DAG a un nœud terminal (\emph{sink}) unique qui est le \emph{genesis vertex}, le sommet genèse. % HOULA sink ? 
%O Maintaining a DAG provides two significant benefits.
L'emploi d'un DAG procure deux avantages :
%O First, it improves efficiency, because a single vote on a DAG vertex implicitly votes for all tran
%O sactions on the path to the genesis vertex.
D'abord, il est plus efficient car un vote sur un sommet vote implicitement pour toutes les transactions sur le chemin vers le sommet genèse.
%O Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
Ensuite, il améliore également la sécurité parce que le entrelace le sort des transactions, similaire en cela à la blockchain de Bitcoin.
%O This renders past decisions difficult to undo without the approval of correct nodes.
Cela rend les décisions passées difficiles à défaire sans l'approbation des nœuds corrects.

% Avalanche's DAG embodies all the transactions that have been proposed to the system, where each transaction is represented as a vertex.
% The DAG is rooted at a protocol-defined, well-known \emph{genesis} vertex.
%O When a client creates a transaction, it names one or more \emph{parents}, which are included inseparably in the transaction and form the edges of the DAG\@.
Quand un client crée une transaction, il nomme un ou plusieurs \emph{parents}, et ceux-ci sont inclus de façon inséparable dans la transaction et forment les bords du DAG\@.
%O The parent-child relationships encoded in the DAG may, but do not need to, correspond to application-specific dependencies; for instance, a child transaction need not spend or have any relationship with the funds received in the parent transaction.
Les relations parent-enfant codées dans le DAG peuvent, sans obligation, correspondre à des dépendances spécifiques aux applications ; par exemple, la transaction d'un enfant ne dépense pas nécessairement, ni n'a forcément de relation avec les fonds reçus lors de la transaction parent.
% These additional relationships entangle the fate of previous decisions made by the system.
%O We use the term \emph{ancestor set} to refer to all transactions reachable via parent edges back in history, and \emph{progeny} to refer to all children transactions and their offspring.
Nous utilisons le terme \emph{ensemble ancêtre} pour nous référer à toutes les transactions accessibles par les arêtes parent historisées, et \emph{progéniture} pour nous référer à toutes les transactions enfant et leur descendance.

%O The central challenge in the maintenance of the DAG is to choose among \emph{conflicting transactions}.
Le principal défi de la maintenance du DAG est de choisir parmi des \emph{trasactions conflictuelles}.
%O The notion of conflict is application-defined and transitive, forming an equivalence relation.
La notion de conflit est définie par l'application et est transitive, formant une relation d'équivalence.
%O In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}

%(Figure~\tronly{\ref{fig:dag-conflict-set}}{\ref{fig:dag-cd}})
(shaded regions in Figure~\ref{fig:dag-cd}), out of which only a single one can be accepted.
Note that the conflict set of a virtuous transaction is always a singleton.
%Note that the graph structure of transactions that spend depend on each other, also known as the UTXO graph, is completely independent of the DAG that Avalanche maintains. As a result, any two vertices may be in conflict. Figure~\ref{fig:dag-conflicting-set} shows an example.

%\tronly{
%\begin{figure}\begin{center}
%    \input{figures/dag-conflict-set.tex}
%    \captionof{figure}{DAG vertices partitioned by conflict sets. At most one vertex in each shaded region will be accepted.}\label{fig:dag-conflict-set}
%\end{center}
%\end{figure}
%}{}

Avalanche instantiates a Snowball instance for each conflict set.
Whereas Snowball uses repeated queries and multiple counters to capture the amount of confidence built in conflicting transactions (colors),
Avalanche takes advantage of the DAG structure and uses a transaction's progeny.
Specifically, when a transaction $T$ is queried, all transactions reachable from $T$ by following the DAG edges are implicitly part of the query.
A node will only respond positively to the query if $T$ and its entire ancestry are currently the preferred option in their respective conflict sets.
If more than a threshold of responders vote positively, the transaction is said to collect a \emph{chit}.
Nodes then compute their \emph{confidence} as the total number of chits in the progeny of that transaction.
They query a transaction just once and rely on new vertices and possible chits, added to the progeny, to build up their confidence.
Ties are broken by an initial preference for first-seen transactions.
Note that chits are decoupled from the DAG structure, making the protocol immune to attacks where
the attacker generates large, padded subgraphs.

\subsection{Avalanche: Specification}%\tronly{}{\vspace{-0.5em}}
\label{subsection:specification}

\newcommand{\codeedges}{\mathit{edges}}
\newcommand{\codedata}{\mathit{data}}
\begin{figure}
\begin{center}
\small
\begin{algorithmic}[1]
    \Procedure{init}{}
        \State $\mathcal{T} \assign \emptyset$ \hspace{1ex}\textrm{// the set of known transactions}
        \State $\mathcal{Q} \assign \emptyset$ \hspace{1ex}\textrm{// the set of queried transactions}
    \EndProcedure
    \Procedure{onGenerateTx}{$\codedata$}
        \State $\codeedges \assign \{T' \gets T: T' \in \Call{parentSelection}{\mathcal{T}}\}$
        \State $T \assign \Call{Tx}{\codedata, \codeedges}$
        \State \Call{onReceiveTx}{$T$}
    \EndProcedure
    \Procedure{onReceiveTx}{$T$}
        \If{$T \notin \mathcal{T}$}
            \If{$\mathcal{P}_T = \emptyset$}
                \State $\mathcal{P}_T \assign \{T\}$, $\mathcal{P}_T\mathit{.pref} \assign T$
                \State $\mathcal{P}_T\mathit{.last} \assign T, \mathcal{P}_T\mathit{.cnt} \assign 0$
            \Else$\ \mathcal{P}_T \assign \mathcal{P}_T \cup \{T\}$
            \EndIf
            \State $\mathcal{T} \assign \mathcal{T} \cup \{T\}$, $c_T \assign 0$.
        \EndIf
    \EndProcedure
    \captionof{figure}{Avalanche: transaction generation.}\label{fig:gossipchain-ongen}
\end{algorithmic}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\small
\begin{algorithmic}[1]
    \Procedure{avalancheLoop}{}
        \While {\codetrue}
            \State\textrm{find  $T$ that satisfies }%\vspace*{-.6\baselineskip}
            $T \in \mathcal{T} \land T \notin \mathcal{Q}$
            %\begin{align*}\hspace{2ex}
            %    T \in \mathcal{T} &\land T \notin \mathcal{Q} \\
            %            &\land (\forall T', T' \gets T: T' \in \mathcal{Q})
            %\end{align*}
            \State $\mathcal{K} \assign \Call{sample}{\mathcal{N}\backslash u, k}$
            \State $P \assign \sum_{v \in \mathcal{K}}\Call{query}{v, T}$
            \If{$P \ge \alpha$}
                \State $c_T \assign 1$
            \State\textrm{// update the preference for ancestors}
            \For{$T' \in \mathcal{T} : T' \stackrel{*}{\gets} T$}
                \If{$d(T') > d(\mathcal{P}_{T'}\mathit{.pref})$}
                    \State $\mathcal{P}_{T'}\mathit{.pref} \assign T'$
                \EndIf
                \If{$T'\neq \mathcal{P}_{T'}\mathit{.last}$}
                    \State $\mathcal{P}_{T'}\mathit{.last} \assign T'$, $\mathcal{P}_{T'}\mathit{.cnt} \assign 1$
                \Else
                    \State \texttt{++}$\mathcal{P}_{T'}\mathit{.cnt}$
                \EndIf
            \EndFor
            \Else
            \For{$T' \in \mathcal{T} : T' \stackrel{*}{\gets} T$}
                    \State$\mathcal{P}_{T'}\mathit{.cnt} \assign 0$
            \EndFor
            \EndIf
            \State\textrm{// otherwise, }$c_T$\textrm{ remains 0 forever}
            \State $\mathcal{Q} \assign \mathcal{Q} \cup \{T\}$ \hspace {1ex} \textrm{// mark T as queried}
        \EndWhile
    \EndProcedure
    \captionof{figure}{Avalanche: the main loop.}\label{fig:gossipchain-main}
\end{algorithmic}
\end{center}
\end{figure}

% TODO IMPORTANT The isPreferred must imply that T is queried, but current writing can be not queried. Also, do you do max over only queried transactions?
\begin{figure}[t]
\begin{center}
\small
\begin{algorithmic}[1]
    \Function{isPreferred}{$T$}
        %\State \Return $d(T) = \max_{T' \in \mathcal{P}_T} d(T')$
        \State \Return $T = \mathcal{P}_T\mathit{.pref}$
    \EndFunction
    \Function{isStronglyPreferred}{$T$}
        \State \Return $\forall T'\in\mathcal{T}, T' \stackrel{*}{\gets} T: \Call{isPreferred}{T'}$
    \EndFunction
    \Function{isAccepted}{$T$}
        \State\Return
            \vspace*{-.5\baselineskip}
        \begin{align*}
            (&(\forall T' \in \mathcal{T}, T' \gets T: \Call{isAccepted}{T'}) \\
                &\land |\mathcal{P}_T| = 1 \land \mathcal{P}_T\mathit{.cnt} \ge \beta_1) \texttt{\hspace{.1in}// safe early commitment} \\
            %& \left(\frac{d(T)}{d'} > \gamma \land d' \ge \beta_2\right)
            \lor &(%\mathcal{P}_T\textrm{.pref} = \mathcal{P}_T\textrm{.last} \land
            \mathcal{P}_T\mathit{.cnt} \ge \beta_2)\texttt{\hspace{.1in}// consecutive counter}
        \end{align*}
    \EndFunction
    \Procedure{onQuery}{$j, T$}
        \State \Call{onReceiveTx}{$T$}
        \State \Call{respond}{$j, \textsc{isStronglyPreferred}(T)$}
    \EndProcedure
    \captionof{figure}{Avalanche: voting and decision primitives.}\label{fig:gossipchain-onquery}
\end{algorithmic}
\end{center}
\end{figure}

Each correct node $u$
keeps track of all transactions it has learned about in set $\mathcal{T}_u$,
partitioned into mutually exclusive conflict sets $\mathcal{P}_T$, $T \in \mathcal{T}_u$.
% Transactions are determined to be conflicting based on a deterministic function known to every node.
Since conflicts are transitive, if $T_i$ and $T_j$ are conflicting, then they belong to the same conflict set, i.e. $\mathcal{P}_{T_i} = \mathcal{P}_{T_j}$. This relation may sound counter-intuitive: conflicting transitions have the \emph{equivalence} relation, because they are equivocations spending the \emph{same} funds.
%\Jon{By $=$ surely you mean $\neq$?}

We write $T' \gets T$ if $T$ has a parent edge to transaction $T'$,
The ``$\stackrel{*}{\gets}$''-relation is its reflexive transitive closure, indicating a path from $T$ to $T'$.
DAGs built by different nodes are guaranteed to be compatible, though at any one time, the two nodes may not have a complete view of all vertices in the system.
Specifically, if $T' \gets T$, then every node in the system that has $T$ will also have $T'$ and the same relation $T' \gets T$; and conversely, if $T' \cancel{\gets} T$, then no nodes will end up with $T' \gets T$.

Each node $u$ can compute a confidence value, $d_u(T)$, from the progeny as follows:
\[ d_u(T) = \sum_{T' \in \mathcal{T}_u, T \stackrel{*}{\gets} T'}c_{uT'}\]
where $c_{uT'}$ stands for the chit value of $T'$ for node $u$. Each transaction initially has a chit value of $0$ before the node gets
the query results. If the node collects a threshold of $\alpha$ yes-votes after the query, the value $c_{uT'}$ is set to 1, otherwise remains $0$ forever.
Therefore, a chit value reflects the result from the one-time query of its associated transaction and becomes immutable afterwards, while $d(T)$ can increase as the DAG grows by collecting more chits in its progeny.
Because $c_T \in \{0, 1\}$, confidence values are monotonic.
%\Jon{the notation $c_{uT}$ shows up before it's definet. I guess it's the number of chits counted for $T$ by participant $u$?}

In addition, node $u$ maintains its own local list of known nodes $\mathcal{N}_u \subseteq \mathcal{N}$ that comprise the system.
For simplicity, we assume for now $\mathcal{N}_u = \mathcal{N}$, and elide subscript $u$ in contexts without ambiguity.
%
% TODO talk to ted and ask how to rephrase this. "the data structure" is unclear, and so is immutability
% This is because a transaction is immutable in the sense that its identity is uniquely determined by both the application data and parent links it embodies.
%Here, notation ``$T' \gets T$'' means $T'$ is one of $T$'s immediate parents, and
%``$\stackrel{*}{\gets}$''-relation is the reflexive transitive closure of ``$\gets$''-relation; namely,
%if $T' \stackrel{*}{\gets} T$, there exists a path from $T$ to $T'$. Specially, ``$T \stackrel{*}{\gets} T$''.

Each node implements an event-driven state machine, centered around a query that serves both to solicit votes on each transaction and to notify other nodes of the existence of newly discovered transactions.
In particular, when node $u$ discovers a transaction $T$ through a query, it starts a one-time query process by sampling $k$ random peers and sending a message to them, after $T$ is delivered via $\textsc{onReceiveTx}$.
% determining the set of conflicting transactions $\mathcal{P}_T$,

Node $u$ answers a query by checking whether each $T'$ such that $T' \stackrel{*}{\gets} T$ is currently preferred among competing transactions $\forall T'' \in \mathcal{P}_{T'}$.
If every single ancestor $T'$ fulfills this criterion, the transaction is said to be \emph{strongly preferred}, and receives a yes-vote (1). A failure of this criterion at any $T'$ yields a no-vote (0).
When $u$ accumulates $k$ responses, it checks whether there are $\alpha$ yes-votes for $T$, and if so grants the chit (chit value $c_T \assign 1$) for $T$.
The above process will yield a labeling of the DAG with a chit value and associated confidence for each transaction~$T$.

\begin{figure}
\begin{center}
    %\tronly{\input{figures/dag-cd}}{\input{figures/dag-cd-compact}}
    \input{figures/dag-cd-compact}
    \captionof{figure}{Example of $\langle \textrm{chit}, \textrm{confidence}\rangle$ values.  Darker boxes indicate transactions with higher confidence values. At most one transaction in each shaded region will be accepted.}
    \label{fig:dag-cd}
\end{center}
\end{figure}

Figure~\ref{fig:dag-cd} illustrates a sample DAG built by Avalanche.
Similar to Snowball, sampling in Avalanche will create a positive feedback for the preference of a single transaction in its conflict set.
For example, because $T_2$ has larger confidence than $T_3$, its descendants are more likely collect chits in the future compared to $T_3$.

\tronly{Similar to Bitcoin, Avalanche leaves determining the acceptance point of a transaction to the application. An application supplies an \textsc{isAccepted} predicate that can take into account the value at risk in the transaction and the chances of a decision being reverted to determine when to decide.}{}

Committing a transaction can be performed through a \emph{safe early commitment}. For virtuous transactions, $T$ is accepted when it is the only transaction in its conflict set and has a confidence not less than threshold $\beta_1$.
As in Snowball, $T$ can also be accepted after a $\beta_2$ number of consecutive successful queries.
If a virtuous transaction fails to get accepted due to a problem
with parents, it could be accepted if reissued with different parents.
Figure~\ref{fig:gossipchain-ongen} shows how Avalanche entangles transactions. Because transactions that consume and generate the same UTXO do not conflict with each other, any transaction can be reissued with different parents.

\tronly{
Figure~\ref{fig:gossipchain-main} illustrates the protocol main loop
executed by each node.
In each iteration, the node attempts to select a transaction $T$ that has not
yet been queried.  If no such
transaction exists, the loop will stall until a new transaction is
added to $\mathcal{T}$.
It then selects $k$ peers and queries those peers.
If more than $\alpha$ of those peers return a positive response, the chit value is set to~1.
After that, it updates the preferred transaction of each conflict set
of the transactions in its ancestry.
Next, $T$ is added to the set $\mathcal{Q}$
so it will never be queried again by the node.
The code that selects additional peers if some of the $k$ peers are
unresponsive is omitted for simplicity.

Figure~\ref{fig:gossipchain-onquery} shows what happens when a node
receives a query for transaction $T$ from peer $j$.
First it adds $T$ to $\mathcal{T}$, unless it already has it.
Then it determines if $T$ is currently strongly preferred.
If so, the node returns a positive response to peer $j$.
Otherwise, it returns a negative response.
Notice that in the pseudocode, we assume when a node knows $T$, it also
recursively knows the entire ancestry of $T$. This can be achieved by
postponing the delivery of $T$ until its entire ancestry is recursively
fetched.
In practice, an additional gossip process that disseminates
transactions is used in parallel, but is not shown in pseudocode for simplicity.
}{}